### 2019-02-25 R Files/Model Writeup.Rmd
Added analysis to Git

---
title: "Modeling Writeup"
author: "Aidan Mullan"
date: "2/10/2019"
header-includes:
  - \usepackage{float}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = 'htb!')
```

# Regression Models

3 models were fit to the Pulmonary Embolism dataset. First, a logistic regression model was fit to determine what factors affect the positivity rate of CT scans for PE. Then, two linear models were fit to the effective dose of radiation and amount of contrast given to patients for their scan.

## Positivity Model

The full model for rate of positive CT scans considered Gender, Age Range, BMI Range, Type of Procedure, Location of Admittance, and Effective Dose as variables of interest. Within the Type of Procedure variable, the 43 "double rule out" conditions had to be removed from analysis since none of these patients had positive test results. In the end, only location and BMI range were found to be significant predictors for positivity. Coefficients and standard errors for this final model are given in Table 1. The default group for BMI is the under 25 group, and the default for location is ED.

\begin{table}[H] \centering 
  \caption{Positivity Model Coefficients} 
\begin{tabular}{p{2cm}|p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & Estimate & Std. Error & p  \\
\hline \\[-1.8ex] 
 Intercept   & $-$2.414 & 0.140 & <.001 $***$ \\
 BMI 25-40   & 0.294 & 0.149 & .048 $*$\\
 BMI $>$40   & 0.107 & 0.204 & .600\\
 Loc: ICU    & 0.543 & 0.172 & .002 $**$\\
 Loc: IN    & 0.295 & 0.151 & .051 $.$\\
 Loc: OUT     & $-$0.147 & 0.262 & .574\\
\hline 
\hline \\[-1.8ex]
\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .005, $***$ < .001}
 \end{tabular}
\end{table}

It is important to note here that although the BMI and Location variables were significant predictors for positivity rate, not all pairwise differences within each factor are significant. Post-hoc analyses were performed using a Bonferroni correction to reduce the chance of spurious significance.

For the BMI groups, a significant difference was found between the under 25 and between 25-40 factors (Z = 1.980, p = .024). No other differences were significant (over 40 and under 25: Z = 0.525, p = .300; over 40 and between 25 and 40: Z = 1.037, p = 0.15)

For the location variable, significant differences were found between in-patients and ED patients (Z = 1.951, p = .026), ICU patients and out-patients (Z = 2.396, p = .008), and between ICU patients and ED patients (Z = 3.150, p < .001). No other pairwise differences were significant (OUT and ED: Z = 0.94, p = .174; OUT and IN: Z = 0.36, p = .360; ICU and IN: Z = 1.288, p = .099).

### Model Interpretation

Since the positivity model utilized logistic regression, the coefficients are interpretated in terms of odds ratios and probability of a positive test result. The way of reading these coefficients is to add the terms of interest and then exponentiate. This provides the odds of a positive test result. For example, the odds that a person with a BMI over 40 who was admitted in the ICU has a positive test result is $$\exp(-2.414+0.107+0.543) = 0.171$$
To convert this to a probability, you only need to compute $\frac{\textrm{Odds}}{1+\textrm{Odds}}$. In the above example, the probability of a positive test result is 
$$\frac{0.171}{1 + 0.171} = .146$$
Therefore we get that the probability of a positive test result for an individual from the ICU with a BMI over 40 is 14.6%.

To compare two groups, such as in-patients and out-patiends, you simply need to divide the two odds. This gives us the odds ratio for a positive test result. The odds ratio for in-patients compared to out-patients is given by 
$$\frac{\exp(.295)}{\exp(-0.147)} = 1.56$$
This tells us that the odds that an in-patient has a positive test result are 56% greater than the odds that an out-patient has a positive test result, assuming both patients fall in the same BMI range. To get the reverse odds ratio (out-patient compared to in-patient), you take the inverse of the original ratio ($\frac{1}{1.56} = 0.64$).

All coefficient interpretations are given below.

#### BMI Groups

\begin{itemize}
  \item 25-40 compared to <25 \\
The odds that an individual with a BMI between 25-40 has a positive test result are 34.2\% greater (OR = 1.342) than the odds for someone under 25 BMI.

  \item >40 compared to <25 \\
The odds that an individual with a BMI over 40 has a positive test result are 11.3\% greater (OR = 1.113) than the odds for someone under 25 BMI.

  \item >40 compared to 25-40 \\
The odds that an individual with a BMI over 40 has a positive test result are 11.1\% less (OR = 0.889) than the odds for someone between 25 and 40 BMI. 
\end{itemize}

#### Location Groups

\begin{itemize}
  \item OUT compared to ED \\
The odds that an individual addmitted as an out-patient has a positive test result are 13.7\% less (OR = 0.863) than the odds for someone addmitted from ED

  \item IN compared to ED \\
The odds that an individual addmitted as an in-patient has a positive test result are 34.3\% greater (OR = 1.343) than the odds for someone addmitted from ED

  \item ICU compared to ED \\
The odds that an individual addmitted in the ICU has a positive test result are 72.1\% greater (OR = 1.721) than the odds for someone addmitted from ED

  \item IN compared to OUT \\
The odds that an individual addmitted as an in-patient has a positive test result are 55.6\% greater (OR = 1.556) than the odds for someone addmitted as an out-patient

  \item ICU compared to OUT \\
The odds that an individual addmitted to the ICU has a positive test result are 99.4\% greater (OR = 1.994) than the odds for someone addmitted as an out-patient

  \item ICU compared to IN \\
The odds that an individual addmitted to the ICU has a positive test result are 28.1\% greater (OR = 1.281) than the odds for someone addmitted from as an in-patient
\end{itemize}
\newpage

## Effective Dosage Model

The full model for effective dose considered CT type, Age and BMI range, Gender, and Location of admittance. Effective dosage was log-transformed to normalize the distribution, and then a linear model was fit to the data. Variable selection found that every variable was significant, and coefficients are given in Table 2. The default factors were: BMI range - under 25, Location - ED, CT type - 128SSwIR, Age - under 18.

\begin{table}[H] \centering 
  \caption{Positivity Model Coefficients} 
\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & Estimate & Std. Error & p  \\
\hline \\[-1.8ex] 
 Intercept   & 0.305 & 0.138 & .027 $*$ \\
 BMI 25-40   & 0.491 & 0.029 & <.001 $***$\\
 BMI $>$40   & 1.131 & 0.040 & <.001 $***$\\
 Loc: ICU    & 0.297 & 0.039 & <.001 $***$\\
 Loc: IN     & 0.013 & 0.032 & .678 \\
 Loc: OUT    & 0.032 & 0.051 & .528\\
 CT: 64SSwIR & 0.109 & 0.064 & .087 $.$ \\
 CT: 64SSWoIR & 0.526 & 0.085 & <.001 $***$ \\
 CT: DS Scanner &$-$0.118 & 0.027 & <.001 $***$ \\
 Age: 18-35 & 0.751 & 0.134 & <.001 $***$\\
 Age: $>$35 & 0.795 & 0.134 & <.001 $***$ \\
 Gender: Male & 0.217 & 0.026 & <.001 $***$ \\
\hline 
\hline \\[-1.8ex]
\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
 \end{tabular}
\end{table}

### Model Interpretation

Since this is a linear model with a log-transformed response variable, the interpretation is similar to the logistic case. All we need to do is add up the coefficients for each factor we are considering and exponentiate. For example, suppose we have a male between the ages of 18 and 35, with a BMI over 40. He was admitted from the ICU and received a CT scan on the 64SSwIR machine. According to our model, his effective dose would be 
$$ \exp(0.305 + 0.217 + 0.751 + 0.297 + 0.109) = 5.36 $$

If we are interested in comparing the change in effective dose between two levels of a single factor, we would do the same thing as the logistic case. Here, we would exponentiate the two coefficients and take their difference. For example, suppose we want to compare the 64SSwIR scanner with the 64SSWoIR scanner. Then we calculate
$$ \frac{\exp(.109)}{\exp(.526)} = .659$$
which tells us that the 64SSwIR scanner is associated with 34.1% decreased effective dose. 
\newpage

## Contrast Model

Similar to the effective dose model, the full model for contrast considered CT type, Age and BMI range, Gender, and Location of admittance. No transformation was necessary, so a standard linear model was fit to the data. Again, variable selection found that every variable was significant, and coefficients are given in Table 3. The default factors were: BMI range - under 25, Location - ED, CT type - 128SSwIR, Age - under 18.

\begin{table}[H] \centering 
  \caption{Positivity Model Coefficients} 
\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & Estimate & Std. Error & p  \\
\hline \\[-1.8ex] 
 Intercept   & 66.364 & 3.184 & <.001 $***$ \\
 BMI 25-40   & 3.343 & 0.805 & <.001 $***$\\
 BMI $>$40   & 17.264 & 1.099 & <.001 $***$\\
 Loc: ICU    & 5.918 & 1.085 & <.001 $***$\\
 Loc: IN     & 1.238 & 0.884 & .161 \\
 Loc: OUT    & 2.696 & 1.420 & .058 $.$\\
 CT: 64SSwIR & 7.507 & 1.767 & <.001 $***$\\
 CT: 64SSWoIR & 6.791 & 2.347 & .004 $**$ \\
 CT: DS Scanner & $-$9.073 & 0.740 & <.001 $***$\\
 Age: 18-35 & 21.716 & 3.849 & <.001 $***$\\
 Age: $>$35 & 21.373 & 3.725 & <.001 $***$ \\
 Gender: Male & 4.324 & 0.707 & <.001 $***$ \\
\hline 
\hline \\[-1.8ex]
\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
 \end{tabular}
\end{table}

### Model Interpretation

This contrast model is a standard linear regression, so the interpretation is pretty straightforward. Since all of our variables are factors, the coefficient for a given factor is the associated difference in contrast between the given factor level and the default. For example, patients with a BMI over 40 are associated with an increase in contrast of 17.26 units over patients with a BMI under 25.

To compare two factor levels that are not the default, we need to take the difference in their coefficients. Suppose we want to compare ICU patients with in-patients. A patient in the ICU receives on average of $5.918-1.238 = 4.68$ units more contrast than an in-patient.

Otherwise, to determine the amount of contrast associated with a given patient, all we need to do is add up the coefficients for each category that the given patient falls into. A 45 year old female with a BMI between 25 and 40 admitted as an out-patient and scanned on a DS scanner would be associated with $66.364+21.373+3.343+2.696-9.073 = 84.703$ units of contrast.

### 2019-02-25 R Files/PE_analysis .Rmd
Added analysis to Git

---
title: "PE_analysis"
author: "Aidan Mullan
date: "2/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Reading in data
PE <- read.csv("PE_clean.csv")
PE_noNA <- na.omit(PE)

PE$Contrast.Given <- as.numeric(PE$Contrast.Given)
```

```{r}
#####-----------------
## Summ Stats for Effective Dose and Contrast
#####-----------------

summary(PE$Eff.Dose)
summary(PE$Contrast.Given)

## Looking at Contrast alone: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$Contrast.Given, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
boxplot(PE$Contrast.Given, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))
summary(PE$Contrast.Given)

## Looking at Effective Dose Alone: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$Eff.Dose, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE$Eff.Dose, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))

## Comparing Age and BMI groups for Eff Dose and Contrast
table(PE$Age_range, PE$BMI_range)

aggregate(Eff.Dose~Age_range+BMI_range, data = PE, FUN = "mean")
aggregate(Eff.Dose~Age_range, data = PE, FUN = "mean")
aggregate(Eff.Dose~BMI_range, data = PE, FUN = "mean")

## Contrast given by age group and BMI group
aggregate(Contrast.Given~Age_range+BMI_range, data = PE, FUN = "mean")
aggregate(Contrast.Given~Age_range, data = PE, FUN = "mean")
aggregate(Contrast.Given~BMI_range, data = PE, FUN = "mean")

## Comparing Scanner Type for Eff Dose and Contrast
aggregate(Eff.Dose~CT.Type, data = PE, FUN = "mean")
aggregate(Contrast.Given~CT.Type, data = PE, FUN = "mean")
aggregate(Eff.Dose~CT.Type, data = PE, FUN = "sd")
aggregate(Contrast.Given~CT.Type, data = PE, FUN = "sd")
```

```{r}
#####-------------------
## Summ Stats for Rate of PE
#####-------------------

PE$PE.Num <- ifelse(PE$Pulmonary.Embolism == "NO", 0, 1)
mean(PE$PE.Num)

## Positivity by Procedure Type
aggregate(PE.Num~Procedure, data = PE, FUN = "mean")

## Two-proportion Z test
p1 <- 0.1093; p2 <- 0.1229; n1 <- 2369; n2 <- 301
pp <- (p1*n1 + p2*n2) / (n1 + n2)
z <- (p1-p2)/(sqrt(pp*(1-pp)*(1/n1 + 1/n2)))
pnorm(z)

## Positivity by Age/BMI
mean(PE$PE.Num)
aggregate(PE.Num~Age_range, data = PE, FUN = "mean")
aggregate(PE.Num~BMI_range, data = PE, FUN = "mean")
aggregate(PE.Num~Age_range+BMI_range, data = PE, FUN = "mean")

## Positivity by Location
aggregate(PE.Num~Location, data = PE, FUN = "mean")
table(PE$Location)
```

```{r}
#####-----------------
## Summ Stats for Patients Testing Positive
####------------------

PE_Pos <- subset(PE, subset = PE.Num == 1)

## Type by Location
table(PE_Pos$Type, PE_Pos$Location)
table(PE_Pos$Type)
table(PE_Pos$Location)

## Type by Age
table(PE_Pos$Type, PE_Pos$Age_range)
table(PE_Pos$Age_range)

## Strain by Location
table(PE_Pos$Strain, PE_Pos$Location)
table(PE_Pos$Strain)

## Strain by Age
table(PE_Pos$Strain, PE_Pos$Age_range)
```

```{r}
#####------------------
## Positivity Model
#####------------------

positivity1 <- glm(PE.Num~Gender+Age_range+Eff.Dose+BMI_range+Procedure+CT.Type+Location,
               family = "binomial", data = PE)
summary(positivity1)

positivity2 <- update(positivity1, .~.-CT.Type)
summary(positivity2)
anova(positivity1, positivity2, test = "Chisq")

positivity3 <- update(positivity2, .~.-Age_range)
anova(positivity2, positivity3, test = "Chisq")
summary(positivity3)

positivity4 <- update(positivity3, .~.-Procedure)
anova(positivity3, positivity4, test = "Chisq") 
summary(positivity4)

positivity5 <- update(positivity4, .~.-Eff.Dose)
summary(positivity5)
anova(positivity4, positivity5, test = "Chisq")

positivity6 <- update(positivity5, .~.-Gender)
anova(positivity5, positivity6, test = "Chisq")
summary(positivity6)

positivity7 <- update(positivity6, .~.-BMI_range)
anova(positivity7, positivity6, test = "Chisq")
summary(positivity7)
## Final model has only Location
## Decided to re-model due to "double rule out" not having any positive test results
```

```{r}
## Removing double rule out procedure from analysis due to no positive test results
PE_nodub <- subset(PE, subset = Procedure != "Double rule out")
pos_nodub1 <- glm(PE.Num~Gender+Age_range+Eff.Dose+BMI_range+Procedure+CT.Type+Location,
               family = "binomial", data = PE_nodub)
summary(pos_nodub1)

pos_nodub2 <- update(pos_nodub1, .~.-Procedure-Age_range-Eff.Dose-CT.Type-Gender)
anova(pos_nodub1, pos_nodub2, test = "Chisq") 
summary(pos_nodub2)
## Final model has BMI and Location

```

```{r}
#####----------------
## Effective Dose Model
#####----------------

effdose1 <- lm(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
summary(effdose1)

effdose2 <- update(effdose1, .~.-Location)
anova(effdose1, effdose2)
## Full model is final model
```

```{r}
#####----------------
## Contrast Model
#####----------------

contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
summary(contrast1)
```

### 2019-02-25 R Files/data_cleaning.Rmd
Added analysis to Git

---
title: "Data_Cleaning"
author: "Aidan Mullan"
date: "2/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
## Read in the raw data files
chest <- read_excel("Chest Only.xlsx")
CTA <- read_excel("CTA.xlsx")

## Remove unnecessary variables (descriptions, repeated columns, etc)
chest_short <- chest[1:2413,c(1:6,8:9,11:16)]
CTA_short <- CTA[,c(1:6,8:9,11:16)]

## Combine chest and CTA datasets
names(CTA_short) <- names(chest_short)
PE_raw <- rbind(chest_short, CTA_short)
```

```{r}
## Collapse factors into main categories for 'Procedure' and 'CT Type'
pro_index <- which(PE_raw$Procedure %in% c("Triple rule out", NA))
PE2 <- PE_raw[-pro_index,]
PE2$Procedure <- ifelse(PE2$Procedure == "Chest only", "Chest Only",
                        ifelse(PE2$Procedure %in% c("Chest and Abdomen", "Neck and Chest",
                                                    "Chest, Abdomen and Pelvis"), "CAP", PE2$Procedure))

PE2$`CT Type` <- ifelse(PE2$`CT Type` %in% c("DS_Force", "DS_Drive"), "DS_Scanner", PE2$`CT Type`)

## Give consistency to 'Location', 'Pulmonary Embolism', 'Type', and 'Strain' factor names
PE2$Location <- ifelse(PE2$Location == "Out", "OUT", 
                       ifelse(PE2$Location == "In", "IN", PE2$Location))

PE2$`Pulmonary Embolism` <- ifelse(PE2$`Pulmonary Embolism` == "No", "NO", 
                       ifelse(PE2$`Pulmonary Embolism` == "Yes", "YES", PE2$`Pulmonary Embolism`))

PE2$Type <- ifelse(PE2$Type == "peripheral", "Peripheral", PE2$Type)
PE2$Type <- ifelse(is.na(PE2$Type), "None", PE2$Type)

PE2$Strain <- ifelse(PE2$Strain %in% c('Yes', 'yes', 'YEs'), "YES",
                     ifelse(PE2$Strain == "No", "NO", PE2$Strain))
PE2$Strain <- ifelse(is.na(PE2$Strain), "None", PE2$Strain)

PE2$BMI[c(112,264,416,453,1876,2188,2217,2371,2650)] <- NA
PE2$BMI <- ifelse(PE2$BMI == "None", NA, PE2$BMI)
```

```{r}
## Convert variables to factors with specified levels
PE2$Gender <- as.factor(PE2$Gender); PE2$Procedure <- as.factor(PE2$Procedure)
PE2$`CT Type` <- as.factor(PE2$`CT Type`); PE2$Location <- as.factor(PE2$Location)
PE2$`Pulmonary Embolism` <- as.factor(PE2$`Pulmonary Embolism`)
PE2$Type <- as.factor(PE2$Type); PE2$Strain <- as.factor(PE2$Strain)
PE2$BMI <- as.numeric(PE2$BMI); PE2$`Contrast Given` <- as.numeric(PE2$`Contrast Given`)

## Construct age/BMI grouping variables
PE2$Age_range <- cut(PE2$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
PE2$BMI_range <- cut(PE2$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))
```

```{r}
write.csv(PE2, "PE_clean.csv")
```

### 2019-02-27 R Files/Update_Analysis.Rmd
Began analysis on the updated dataset

---
title: "Analysis2"
author: "Aidan Mullan"
date: "2/26/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
PE <- read_excel("Data/Updates/Book2.xlsx")
PE$BMI <- as.numeric(PE$BMI)
PE <- na.omit(PE)
summary(PE)
```

```{r}

## Looking at Contrast: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$`Contrast Given`, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
boxplot(PE$`Contrast Given`, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given", yaxp = c(0,150,10))
par(mfrow = c(1,1))

## Looking at Effective Dose: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE$`Eff Dose`, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE$`Eff Dose`, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))
```

```{r}
## Plotting Contrast by Age (10 Year Intervals)
age_cuts <- c("<10", "10-20", "20-30", "30-40", "40-50","50-60", "60-70",
              "70-80", "80-90", "90-100", ">100")
PE$Age10 <- cut(PE$Age, breaks = seq(0,110,10), labels = age_cuts)
summary(PE$Age10)

contrast_by_age <- tapply(PE$`Contrast Given`, PE$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Avergae Contrast Administered by Age")
points(contrast_by_age)
axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)

## Plotting Radiation by Age (10 Year Intervals)
rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Mean Effective Dose by Age")
points(rad_by_age)
axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)

## Plotting Contrast by BMI (5 Unit Intervals)
BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
              "40-45", "45-50", "50-55", "55-60", ">60")
PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
summary(PE$BMI10)

contrast_by_bmi <- tapply(PE$`Contrast Given`, PE$BMI10, "mean")
plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
     main = "Avergae Contrast Administered by BMI")
points(contrast_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)

## Plotting Radiation by BMI (5 Unit Intervals)
rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
     main = "Mean Effective Dose by BMI")
points(rad_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)
```

### 2019-02-27 R Files/Updates_Writeup.Rmd
Began analysis on the updated dataset

---
title: "Update_Writeup"
author: "Aidan Mullan"
date: "2/26/2019"
header-includes:
  - \usepackage{float}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'htb!')
```

```{r, echo = FALSE}
library(readxl)
PE <- read_excel("Data/Updates/Book2.xlsx")
PE$BMI <- as.numeric(PE$BMI)
PE <- na.omit(PE)

age_cuts <- c("<20", "20-30", "30-40", "40-50","50-60", "60-70",
              "70-80", "80-90", ">90")
PE$Age10 <- cut(PE$Age, breaks = c(-1,seq(20,90,10), 110), labels = age_cuts)

BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
              "40-45", "45-50", "50-55", "55-60", ">60")
PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
```

## Effective Dose and Contrast

#### Summary Statistics

Numerical summaries of both effective dose and contrast are given in Table 1. Using the fixed data for effective dosage of radiation administered to patients, we retrieve the plots shown in Figure 1. Here we can see the right-skew in the histogram suggesting several patients recieved much higher amounts of radiation than average.

\begin{table}[H] \centering 
  \caption{Summary Statistics of Radiation and Contrast} 
\begin{tabular}{p{3cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 & Min & 1st Q & Median & Mean & 3rd Q & Max \\ 
\hline \\[-1.8ex] 
Effective Dose & 0.10 & 3.06 & 4.30 & 5.631 & 6.36 & 48.83 \\
Contrast       & 5.0 & 75.0 & 95.0 & 88.5 & 98.0 & 150.0 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

```{r, echo = FALSE}
par(mfrow = c(1,2))
hist(PE$`Eff Dose`, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE$`Eff Dose`, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))
summary(PE$`Eff Dose`)
```

We also have constructed an updated histogram and boxplot for the amount of contrast given to patients, which are shown in Figure 2. Here we see a centralized distribution with some extreme values on both sides.

```{r}
par(mfrow = c(1,2))
hist(PE$`Contrast Given`, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
boxplot(PE$`Contrast Given`, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given", yaxp = c(0,150,10))
par(mfrow = c(1,1))
```

We are also interested in the relationship of the effecive dose and contrast administered with both the age and BMI of the patient. Patients were divided into 10-year age groups and 5-unit BMI groups. In both cases, cut-off values were determined based on the number of patients that fell into the extreme groups. For example, there was only one patient in the data under the age of 10, so they were added into the "<20" group. The number of patients that belong to each age group and BMI group is given in Table 2 and 3 respectively. Figure 3 shows the average radiation administered for each of these age groups.

```{r}
summary(PE$Age10)
rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Figure 3: Mean Effective Dose by Age", ylim = c(3,7))
points(rad_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)
```

## Contrast Given

### 2019-02-27 R Files/Updates_Writeup.tex
Began analysis on the updated dataset

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Update\_Writeup},
            pdfauthor={Aidan Mullan},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Update\_Writeup}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Aidan Mullan}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2/26/2019}

\usepackage{float}

\begin{document}
\maketitle

\begin{verbatim}
## Warning in read_fun(path = path, sheet_i = sheet, limits = limits, shim =
## shim, : Expecting numeric in I1195 / R1195C9: got 'l00'
\end{verbatim}

\begin{verbatim}
## Warning: NAs introduced by coercion
\end{verbatim}

\subsection{Effective Dose}\label{effective-dose}

\paragraph{Summary Statistics}\label{summary-statistics}

Using the fixed data for effective dosage of radiation administered to
patients, we retrieve the plots shown in Figure 1. Here we can see the
right-skew in the histogram suggesting several patients recieved much
higher amounts of radiation than average. The mean effective dose given
to a patient was 5.631 millisieverts, with a range of 0.1 to 48.8
millisieverts. Out of all 2382 patients, only 54 recieved a radiation
dose above 20 millisieverts.

\begin{table}[H] \centering 
  \caption{Summary Statistics of Radiation and Contrast} 
\begin{tabular}{p{3cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
\\[-1.8ex] \hline 
\hline
 & Min & 1st Q & Median & Mean & 3rd Q & Max \\ 
\hline \\[-1.8ex] 
Effective Dose & 0.10 & 3.06 & 4.30 & 5.631 & 6.36 & 48.83 \\
Contrast       & 5.0 & 75.0 & 95.0 & 88.5 & 98.0 & 150.0 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-2-1.pdf}

\begin{verbatim}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   0.098   3.055   4.298   5.631   6.356  48.832
\end{verbatim}

We are also interested in the relationship between the effecive dose
administered and both the age and BMI of the patient. For the case of
age, patients were grouped into 10-year age groups. The number of
patients that belong to group is given below. Figure 2 shows the average
radiation administered for each of these age groups.

\begin{verbatim}
##   <20 20-30 30-40 40-50 50-60 60-70 70-80 80-90   >90 
##    55   133   222   294   514   562   397   179    26
\end{verbatim}

\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-3-1.pdf}

\subsection{Contrast Given}\label{contrast-given}

\end{document}

### 2019-02-27 R Files/Updates_Writeup.Rmd
Finished EDA for contrast and effective dose, knitted to pdf

@@
 The number of patients that belong to each age group and BMI group is given in Table 2 and 3 respectively.
-Figure 3 shows the average radiation administered for each of these age groups.
 
-```{r}
-summary(PE$Age10)
+\begin{table}[H] \centering 
+  \caption{Breakdown Patients by Age Group} 
+\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
+\\[-1.8ex] \hline 
+\hline
+ <20 & 20-30 & 30-40 & 40-50& 50-60 & 60-70 & 70-80 & 80-90 & >90 \\ 
+ \hline \\[-1.8ex] 
+55 & 133 & 222 & 294 & 514 & 562 & 397 & 179 & 26 \\
+\hline 
+\hline \\[-1.8ex]
+ \end{tabular}
+\end{table}
+
+\begin{table}[H] \centering 
+  \caption{Breakdown Patients by BMI Group} 
+\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
+\\[-1.8ex] \hline 
+\hline
+ <15 & 15-20 & 20-25 & 25-30 & 30-35 & 35-40 & 40-45 & 45-50 & 50-55 & 55-60 & >60 \\ 
+ \hline \\[-1.8ex] 
+14 & 158 & 486 & 595 & 467 & 276 & 171 & 106 & 56 & 21 & 32 \\
+\hline 
+\hline \\[-1.8ex]
+ \end{tabular}
+\end{table}
+
+Figure 3 shows the average radiation and contrast administered for each of the age groups, and Figure 4 shows the average radiation and contrast administered for the BMI groups.
+
+```{r, fig.height = 5, fig.width = 10}
+par(mfrow = c(1,2))
 rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
 plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
-     main = "Figure 3: Mean Effective Dose by Age", ylim = c(3,7))
+     main = "Figure 3a: Mean Effective Dose by Age", ylim = c(3,7), cex.main = 1)
 points(rad_by_age)
-axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)
+axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
+
+contrast_by_age <- tapply(PE$`Contrast Given`, PE$Age10, "mean")
+plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
+     main = "Figure 3b: Mean Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(75,95))
+points(contrast_by_age)
+axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
+axis(2, at = seq(75,95,5))
+par(mfrow = c(1,1))
 ```
 
-## Contrast Given
+```{r, fig.height = 5, fig.width = 10}
+par(mfrow = c(1,2))
+rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
+plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
+     main = "Figure 4a: Mean Effective Dose by BMI", cex.main = 1)
+points(rad_by_bmi)
+axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
+
+contrast_by_bmi <- tapply(PE$`Contrast Given`, PE$BMI10, "mean")
+plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
+     main = "Figure 4b: Mean Contrast Administered by BMI", cex.main = 1)
+points(contrast_by_bmi)
+axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
+par(mfrow = c(1,1))
+```
+
+Interestingly, we see nearly identical patterns for both the radiation and contrast administered when divided into either age groups or BMI groups.
+For age, we see a nearly quadratic relationship, with more radiation and contrast being administered for groups between 30 and 60 years of age, and lower amounts administered to both younger and older groups.
+
+On the other hand, we have seem to have a purely increasing relationship for radiation and contrast with BMI.
+As a patient's BMI increases, we can see that the average effective dosage and amount of contrast given both increase.
+
+From these results, it appears that the amount of radiation and contrast administered to a patient are correlated.
+Indeed, we find a Pearson's correlation of 0.47, which is significant (t(2380) = 26.05, p < .001).

### 2019-02-27 R Files/Updates_Writeup.tex
Finished EDA for contrast and effective dose, knitted to pdf

@@
 
-\subsection{Effective Dose}\label{effective-dose}
+\subsection{Effective Dose and
+Contrast}\label{effective-dose-and-contrast}
 
@@
 
-Using the fixed data for effective dosage of radiation administered to
-patients, we retrieve the plots shown in Figure 1. Here we can see the
-right-skew in the histogram suggesting several patients recieved much
-higher amounts of radiation than average. The mean effective dose given
-to a patient was 5.631 millisieverts, with a range of 0.1 to 48.8
-millisieverts. Out of all 2382 patients, only 54 recieved a radiation
-dose above 20 millisieverts.
+Numerical summaries of both effective dose and contrast are given in
+Table 1. Using the fixed data for effective dosage of radiation
+administered to patients, we retrieve the plots shown in Figure 1. Here
+we can see the right-skew in the histogram suggesting several patients
+recieved much higher amounts of radiation than average.
 
@@
 
-We are also interested in the relationship between the effecive dose
-administered and both the age and BMI of the patient. For the case of
-age, patients were grouped into 10-year age groups. The number of
-patients that belong to group is given below. Figure 2 shows the average
-radiation administered for each of these age groups.
-
-\begin{verbatim}
-##   <20 20-30 30-40 40-50 50-60 60-70 70-80 80-90   >90 
-##    55   133   222   294   514   562   397   179    26
-\end{verbatim}
+We also have constructed an updated histogram and boxplot for the amount
+of contrast given to patients, which are shown in Figure 2. Here we see
+a centralized distribution with some extreme values on both sides.
 
 \includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+We are also interested in the relationship of the effecive dose and
+contrast administered with both the age and BMI of the patient. Patients
+were divided into 10-year age groups and 5-unit BMI groups. In both
+cases, cut-off values were determined based on the number of patients
+that fell into the extreme groups. For example, there was only one
+patient in the data under the age of 10, so they were added into the
+``\textless{}20'' group. The number of patients that belong to each age
+group and BMI group is given in Table 2 and 3 respectively.
+
+\begin{table}[H] \centering 
+  \caption{Breakdown Patients by Age Group} 
+\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
+\\[-1.8ex] \hline 
+\hline
+ <20 & 20-30 & 30-40 & 40-50& 50-60 & 60-70 & 70-80 & 80-90 & >90 \\ 
+ \hline \\[-1.8ex] 
+55 & 133 & 222 & 294 & 514 & 562 & 397 & 179 & 26 \\
+\hline 
+\hline \\[-1.8ex]
+ \end{tabular}
+\end{table}
+
+\begin{table}[H] \centering 
+  \caption{Breakdown Patients by BMI Group} 
+\begin{tabular}{p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}}
+\\[-1.8ex] \hline 
+\hline
+ <15 & 15-20 & 20-25 & 25-30 & 30-35 & 35-40 & 40-45 & 45-50 & 50-55 & 55-60 & >60 \\ 
+ \hline \\[-1.8ex] 
+14 & 158 & 486 & 595 & 467 & 276 & 171 & 106 & 56 & 21 & 32 \\
+\hline 
+\hline \\[-1.8ex]
+ \end{tabular}
+\end{table}
+
+Figure 3 shows the average radiation and contrast administered for each
+of the age groups, and Figure 4 shows the average radiation and contrast
+administered for the BMI groups.
+
+\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+\includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-5-1.pdf}
 

### 2019-02-27 R Files/Update_Analysis.Rmd
Added updated data, began model re-fitting

@@
 ```{r}
-library(readxl)
-PE <- read_excel("Data/Updates/Book2.xlsx")
-PE$BMI <- as.numeric(PE$BMI)
-PE <- na.omit(PE)
-summary(PE)
+PE <- read.csv("Data/Updates/PE_update_clean.csv")[,-1]
 ```
@@
 par(mfrow = c(1,2))
-hist(PE$`Contrast Given`, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
+hist(PE$Contrast.Given, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
      xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
-boxplot(PE$`Contrast Given`, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
+boxplot(PE$Contrast.Given, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
         ylab = "Contrast Given", yaxp = c(0,150,10))
@@
 par(mfrow = c(1,2))
-hist(PE$`Eff Dose`, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
+hist(PE$Eff.Dose, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
      xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
 axis(side = 1, at = seq(0,100,10))
-boxplot(PE$`Eff Dose`, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
+boxplot(PE$Eff.Dose, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
         ylab = "Effective Dose")
@@
 
-contrast_by_age <- tapply(PE$`Contrast Given`, PE$Age10, "mean")
+contrast_by_age <- tapply(PE$Contrast.Given, PE$Age10, "mean")
 plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
@@
 ## Plotting Radiation by Age (10 Year Intervals)
-rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
+rad_by_age <- tapply(PE$Eff.Dose, PE$Age10, "mean")
 plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
@@
 
-contrast_by_bmi <- tapply(PE$`Contrast Given`, PE$BMI10, "mean")
+contrast_by_bmi <- tapply(PE$Contrast.Given, PE$BMI10, "mean")
 plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
@@
 ## Plotting Radiation by BMI (5 Unit Intervals)
-rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
+rad_by_bmi <- tapply(PE$Eff.Dose, PE$BMI10, "mean")
 plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
@@
 ```
+
+#####----------------------------------------------------
+
+#####----------------------------------------------------
+
+#####----------------------------------------------------
+
+```{r}
+PE$Age_range <- cut(PE$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
+PE$BMI_range <- cut(PE$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))
+
+effdose1 <- lm(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
+```
+
+```{r}
+contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
+```

### 2019-02-27 R Files/Update_Cleaning.Rmd
Added updated data, began model re-fitting

---
title: "Update_Cleaning"
author: "Aidan Mullan"
date: "2/27/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
PE_raw <- read_excel("Data/Updates/chest_only.xlsx")
names(PE_raw)

PE_short <- PE_raw[1:2413,c(3,4,6,8,9,11,12,13)]
unique(PE_short$Location)

PE_short$Location <- ifelse(PE_short$Location == "Out", "OUT", 
                       ifelse(PE_short$Location == "In", "IN", PE_short$Location))
PE_short$BMI <- as.numeric(PE_short$BMI)
PE_short$`Contrast Given` <- as.numeric(PE_short$`Contrast Given`)
PE_short$`CT Type` <- ifelse(PE_short$`CT Type` %in% c("DS_Force", "DS_Drive"),
                             "DS_Scanner", PE_short$`CT Type`)
PE_short$Procedure <- ifelse(PE_short$Procedure == "Chest only", "Chest Only", PE_short$Procedure)

PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
PE <- na.omit(PE_chest)

write.csv(PE, file = "PE_update_clean.csv")
```

### 2019-02-27 R Files/Updates_Writeup.Rmd
Added updated data, began model re-fitting

@@
 ```{r, echo = FALSE}
-library(readxl)
-PE <- read_excel("Data/Updates/Book2.xlsx")
+PE <- read.csv("Data/Updates/PE_update_clean.csv")
 PE$BMI <- as.numeric(PE$BMI)
@@
 par(mfrow = c(1,1))
-summary(PE$`Eff Dose`)
 ```
@@
 par(mfrow = c(1,2))
-rad_by_age <- tapply(PE$`Eff Dose`, PE$Age10, "mean")
+rad_by_age <- tapply(PE$Eff.Dose, PE$Age10, "mean")
 plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
@@
 par(mfrow = c(1,2))
-rad_by_bmi <- tapply(PE$`Eff Dose`, PE$BMI10, "mean")
+rad_by_bmi <- tapply(PE$Eff.Dose, PE$BMI10, "mean")
 plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",

### 2019-02-27 R Files/Updates_Writeup.tex
Added updated data, began model re-fitting

@@
 
-\begin{verbatim}
-## Warning in read_fun(path = path, sheet_i = sheet, limits = limits, shim =
-## shim, : Expecting numeric in I1195 / R1195C9: got 'l00'
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: NAs introduced by coercion
-\end{verbatim}
-
 \subsection{Effective Dose and
@@
 \includegraphics{Updates_Writeup_files/figure-latex/unnamed-chunk-2-1.pdf}
-
-\begin{verbatim}
-##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-##   0.098   3.055   4.298   5.631   6.356  48.832
-\end{verbatim}
 
@@
 
-\subsection{Contrast Given}\label{contrast-given}
+Interestingly, we see nearly identical patterns for both the radiation
+and contrast administered when divided into either age groups or BMI
+groups. For age, we see a nearly quadratic relationship, with more
+radiation and contrast being administered for groups between 30 and 60
+years of age, and lower amounts administered to both younger and older
+groups.
+
+On the other hand, we have seem to have a purely increasing relationship
+for radiation and contrast with BMI. As a patient's BMI increases, we
+can see that the average effective dosage and amount of contrast given
+both increase.
+
+From these results, it appears that the amount of radiation and contrast
+administered to a patient are correlated. Indeed, we find a Pearson's
+correlation of 0.47, which is significant (t(2380) = 26.05, p
+\textless{} .001).
 

### 2019-03-01 R Files/Updates_Writeup.Rmd
Refit eff dose and contrast models

@@
 plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
-     main = "Figure 4b: Mean Contrast Administered by BMI", cex.main = 1)
+     main = "Figure 4b: Mean Contrast Administered by BMI", cex.main = 1, ylim = c(75,96))
 points(contrast_by_bmi)
 axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
+axis(2, at = seq(75,95,5))
 par(mfrow = c(1,1))
@@
 
-Interestingly, we see nearly identical patterns for both the radiation and contrast administered when divided into either age groups or BMI groups.
-For age, we see a nearly quadratic relationship, with more radiation and contrast being administered for groups between 30 and 60 years of age, and lower amounts administered to both younger and older groups.
+For our age groups, we see a nearly quadratic relationship with effective dose, with more radiation and contrast being administered for groups between 30 and 60 years of age, and lower amounts administered to both younger and older groups.
+There doesn't appear to be any prominent relationship between age and contrast.
 
-On the other hand, we have seem to have a purely increasing relationship for radiation and contrast with BMI.
-As a patient's BMI increases, we can see that the average effective dosage and amount of contrast given both increase.
+Now looking at BMI, there appears to be a linear association between radiation and BMI, with an increase in effective dose as BMI increases.
+Again, there does not appear to be any relationship between contrast and BMI
 
@@
 Indeed, we find a Pearson's correlation of 0.47, which is significant (t(2380) = 26.05, p < .001).
+
+## Effective Dose Model
+
+With the fixed values for effective dose, we refit our linear model.
+Again, a log-transformation was performed to normalize the effective dose variable.
+The initial predictors considered were CT Type, BMI (<25, 25-40, >40), Age (<18, 18-35, >35), Gender, and Location of Admission.
+From the initial model, the only significant location factor was the ICU.
+In this case, we condensed the \textit{Location} variable into two groups: patients admitted in the ICU and other patients.
+No significant difference in predictive accuracy was determined between the model with the full location variable and the model with the condensed location variable, so we prefer the condensed model.
+The final model with coefficients is given in Table 4.
+The baseline factors for each variable are: BMI: <25, Location: ICU, CT: 128SSwIR, Age: <18, Gender: Female.
+
+\begin{table}[H] \centering 
+  \caption{Effective Dosage Model Coefficients} 
+\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
+\\[-1.8ex] \hline 
+\hline
+    & Estimate & Std. Error & p  \\
+\hline \\[-1.8ex] 
+ Intercept   & 0.472 & 0.111 & .027 $***$ \\
+ BMI 25-40   & 0.516 & 0.023 & <.001 $***$\\
+ BMI $>$40   & 1.156 & 0.032 & <.001 $***$\\
+ Loc: Other  & $-$0.106 & 0.032 & <.001 $***$\\
+ CT: 64SSwIR & 0.019 & .049 & .699  \\
+ CT: 64SSWoIR & 0.582 & 0.065 & <.001 $***$ \\
+ CT: DS Scanner &$-$0.203 & 0.021 & <.001 $***$ \\
+ Age: 18-35 & 0.611 & 0.111 & <.001 $***$\\
+ Age: $>$35 & 0.624 & 0.107 & <.001 $***$ \\
+ Gender: Male & 0.227 & 0.020 & <.001 $***$ \\
+\hline 
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
+ \end{tabular}
+\end{table}
+
+#### Comparison of Scanner Type
+
+Post-hoc analysis using Tukey's HSD was conducted for the difference in effective dosage administered based on the type of scanner being used.
+This analysis found that DS scanners were associated with significantly lower radiation levels than all of the other scanners ($p < .001$ for all 3).
+Moreover, the 64SSwIR scanner had significantly higher radiation levels than both the 64SSwoIR and 128SSwIR ($p<.001$ for both).
+Lastly, there was no difference in effective dose between the 64SSwIR and the 128SSwIR scanners.
+
+The following illustrate the differences in effective dose by scanner as determined by our model:
+\begin{itemize}
+\item The DS scanners was associated with 81.6\% of the radiation given for the 128SSwIR scanner and 80.1\% of the radiation used for the 64SSwIR scanner, decreases of 18.4\% and 19.9\% respectively
+\item The DS scanners was associated with 45.6\% of the radiation given for the 64SSwoIR scanner, a decrease of 54.4\%
+\item The 64SSwoIR scanner was associate with a 67.5\% increase in radiation over the 128SSwIR scanner, and a 64.4\% increase in radiation over the 64SSwIR scanner 
+\end{itemize}
+
+To note, since there was no significant difference between the 64SSwIR and 128SSwIR scanners, the first and third bullet points above can be generalized.
+This would tells us that DS scanners use ~19\% less radiation than the IR scanners, and the 64SSwoIR scanners use ~65\% more radiation than the IR scanners.
+
+## Contrast Model
+
+We also refit the contrast model using the updated data set.
+We considered a full model including CT Type, BMI (<25, 25-40, >40), Age (<18, 18-35, >35), Gender, and Location of Admission.
+Model coefficients are given in Table 5.
+Interestingly, with the updated data we no longer have significance for Age groups, BMI groups, or Gender.
+This would suggest that some of the erratic values from the previous dataset that have been removed were strongly linked to specific age, BMI, and gender levels.
+
+\begin{table}[H] \centering 
+  \caption{Contrast Model Coefficients} 
+\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
+\\[-1.8ex] \hline 
+\hline
+    & Estimate & Std. Error & p  \\
+\hline \\[-1.8ex] 
+ Intercept   & 88.580 & 0.592 & .027 $***$ \\
+ CT: 64SSwIR & $-$6.953 & 1.831 & <.001 $***$  \\
+ CT: 64SSWoIR & $-$0.214 & 2.376 & .928 \\
+ CT: DS Scanner &$-$1.799 & 0.751 & .017 $*$ \\
+ Loc: ICU & 3.411 & 1.161 & .003 $**$\\
+ Loc: IN & 1.635 & 0.900 & .069 $.$ \\
+ Loc: OUT & 4.188 & 1.435 & .004 $**$ \\
+\hline 
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
+ \end{tabular}
+\end{table}
+
+#### Comparison of Scanner Type
+
+As with the effective dose model, post hoc analysis using Tukey's HSD for the type of scanner.
+Here we find that the only significant difference was between the 64SSwIR and 64SSwoIR scanners ($p = .022$).
+From our model, this difference equates to the 64SSwIR scanners using 6.95 units less contrast than the 64SSwoIR scanners.
+
+The interesting part of this finding is that our model found a significant effect for DS Scanners over 128SSwIR scanners, but the post-hoc analysis did not.
+Tukey's HSD tends to be more conservative, and so it is more difficult to find significance.
+This is primarily to account for the number of pair-wise comparisons being made, whereas standard linear regression does not need to account for this.

### 2019-03-01 R Files/Updates_Writeup.tex
Refit eff dose and contrast models

@@
 
-Interestingly, we see nearly identical patterns for both the radiation
-and contrast administered when divided into either age groups or BMI
-groups. For age, we see a nearly quadratic relationship, with more
-radiation and contrast being administered for groups between 30 and 60
-years of age, and lower amounts administered to both younger and older
-groups.
+For our age groups, we see a nearly quadratic relationship with
+effective dose, with more radiation and contrast being administered for
+groups between 30 and 60 years of age, and lower amounts administered to
+both younger and older groups. There doesn't appear to be any prominent
+relationship between age and contrast.
 
-On the other hand, we have seem to have a purely increasing relationship
-for radiation and contrast with BMI. As a patient's BMI increases, we
-can see that the average effective dosage and amount of contrast given
-both increase.
+Now looking at BMI, there appears to be a linear association between
+radiation and BMI, with an increase in effective dose as BMI increases.
+Again, there does not appear to be any relationship between contrast and
+BMI
 
@@
 
+\subsection{Effective Dose Model}\label{effective-dose-model}
+
+With the fixed values for effective dose, we refit our linear model.
+Again, a log-transformation was performed to normalize the effective
+dose variable. The initial predictors considered were CT Type, BMI
+(\textless{}25, 25-40, \textgreater{}40), Age (\textless{}18, 18-35,
+\textgreater{}35), Gender, and Location of Admission. From the initial
+model, the only significant location factor was the ICU. In this case,
+we condensed the \textit{Location} variable into two groups: patients
+admitted in the ICU and other patients. No significant difference in
+predictive accuracy was determined between the model with the full
+location variable and the model with the condensed location variable, so
+we prefer the condensed model. The final model with coefficients is
+given in Table 4. The baseline factors for each variable are: BMI:
+\textless{}25, Location: ICU, CT: 128SSwIR, Age: \textless{}18, Gender:
+Female.
+
+\begin{table}[H] \centering 
+  \caption{Effective Dosage Model Coefficients} 
+\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
+\\[-1.8ex] \hline 
+\hline
+    & Estimate & Std. Error & p  \\
+\hline \\[-1.8ex] 
+ Intercept   & 0.472 & 0.111 & .027 $***$ \\
+ BMI 25-40   & 0.516 & 0.023 & <.001 $***$\\
+ BMI $>$40   & 1.156 & 0.032 & <.001 $***$\\
+ Loc: Other  & $-$0.106 & 0.032 & <.001 $***$\\
+ CT: 64SSwIR & 0.019 & .049 & .699  \\
+ CT: 64SSWoIR & 0.582 & 0.065 & <.001 $***$ \\
+ CT: DS Scanner &$-$0.203 & 0.021 & <.001 $***$ \\
+ Age: 18-35 & 0.611 & 0.111 & <.001 $***$\\
+ Age: $>$35 & 0.624 & 0.107 & <.001 $***$ \\
+ Gender: Male & 0.227 & 0.020 & <.001 $***$ \\
+\hline 
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
+ \end{tabular}
+\end{table}
+
+\paragraph{Comparison of Scanner Type}\label{comparison-of-scanner-type}
+
+Post-hoc analysis using Tukey's HSD was conducted for the difference in
+effective dosage administered based on the type of scanner being used.
+This analysis found that DS scanners were associated with significantly
+lower radiation levels than all of the other scanners (\(p < .001\) for
+all 3). Moreover, the 64SSwIR scanner had significantly higher radiation
+levels than both the 64SSwoIR and 128SSwIR (\(p<.001\) for both).
+Lastly, there was no difference in effective dose between the 64SSwIR
+and the 128SSwIR scanners.
+
+The following illustrate the differences in effective dose by scanner as
+determined by our model:
+
+\begin{itemize}
+\item The DS scanners was associated with 81.6\% of the radiation given for the 128SSwIR scanner and 80.1\% of the radiation used for the 64SSwIR scanner, decreases of 18.4\% and 19.9\% respectively
+\item The DS scanners was associated with 45.6\% of the radiation given for the 64SSwoIR scanner, a decrease of 54.4\%
+\item The 64SSwoIR scanner was associate with a 67.5\% increase in radiation over the 128SSwIR scanner, and a 64.4\% increase in radiation over the 64SSwIR scanner 
+\end{itemize}
+
+To note, since there was no significant difference between the 64SSwIR
+and 128SSwIR scanners, the first and third bullet points above can be
+generalized. This would tells us that DS scanners use
+\textasciitilde{}19\% less radiation than the IR scanners, and the
+64SSwoIR scanners use \textasciitilde{}65\% more radiation than the IR
+scanners.
+
+\subsection{Contrast Model}\label{contrast-model}
+
+We also refit the contrast model using the updated data set. We
+considered a full model including CT Type, BMI (\textless{}25, 25-40,
+\textgreater{}40), Age (\textless{}18, 18-35, \textgreater{}35), Gender,
+and Location of Admission. Model coefficients are given in Table 5.
+Interestingly, with the updated data we no longer have significance for
+Age groups, BMI groups, or Gender. This would suggest that some of the
+erratic values from the previous dataset that have been removed were
+strongly linked to specific age, BMI, and gender levels.
+
+\begin{table}[H] \centering 
+  \caption{Contrast Model Coefficients} 
+\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
+\\[-1.8ex] \hline 
+\hline
+    & Estimate & Std. Error & p  \\
+\hline \\[-1.8ex] 
+ Intercept   & 88.580 & 0.592 & .027 $***$ \\
+ CT: 64SSwIR & $-$6.953 & 1.831 & <.001 $***$  \\
+ CT: 64SSWoIR & $-$0.214 & 2.376 & .928 \\
+ CT: DS Scanner &$-$1.799 & 0.751 & .017 $*$ \\
+ Loc: ICU & 3.411 & 1.161 & .003 $**$\\
+ Loc: IN & 1.635 & 0.900 & .069 $.$ \\
+ Loc: OUT & 4.188 & 1.435 & .004 $**$ \\
+\hline 
+\hline \\[-1.8ex]
+\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
+ \end{tabular}
+\end{table}
+
+\paragraph{Comparison of Scanner
+Type}\label{comparison-of-scanner-type-1}
+
+As with the effective dose model, post hoc analysis using Tukey's HSD
+for the type of scanner. Here we find that the only significant
+difference was between the 64SSwIR and 64SSwoIR scanners (\(p = .022\)).
+From our model, this difference equates to the 64SSwIR scanners using
+6.95 units less contrast than the 64SSwoIR scanners.
+
+The interesting part of this finding is that our model found a
+significant effect for DS Scanners over 128SSwIR scanners, but the
+post-hoc analysis did not. Tukey's HSD tends to be more conservative,
+and so it is more difficult to find significance. This is primarily to
+account for the number of pair-wise comparisons being made, whereas
+standard linear regression does not need to account for this.
+
 \end{document}

### 2019-03-03 R Files/Initial Model/Summ_Stat Writeup.Rmd
Added folders to separate analysis

---
title: "PE Summary Statistics"
author: "Aidan Mullan"
date: "2/7/2019"
header-includes:
  - \usepackage{float}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'htb!')
```

```{r}
PE <- read.csv("PE_clean.csv")
PE$Contrast.Given <- as.numeric(PE$Contrast.Given)
```

# Summary Statistics

The first two variables we will look at are the Effective Dose of radiation given to patients and the amount of contrast given before the patient's scan. A focus will be placed on how these two measures differ between age groups (under 18, 18-35, older than 35) and BMI groups (under 25, 25-40, above 40).

\textbf{Note:} Any differences or patterns in the data described here are neither definitively significant nor non-significant unless specifically stated. Significance in variable effect will be determined by later regression.

## Effective Radiation Dose and Contrast Given to Patients

Figure 1 shows the distribution of the dose of radiation given to patients during their scan. The average dose administered was 7.23 sieverts. There are a number of patients receiving a substantially higher dose than this average. 522 patients received more than 10 sieverts, 163 received more than 20, and 32 patients received an effective does of more than 40 sieverts. The largest dose received by any patient was 104.73 sieverts.

```{r, echo = FALSE}
par(mfrow = c(1,2))
hist(PE$Eff.Dose, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE$Eff.Dose, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))

summary(PE$Eff.Dose)
```

The distribution in the amount of contrast given to patients prior to their CT scan appears to have a pseudo-normal shape centered around 90, as seen in Figure 2. The average amount of contrast used for a patient was 91.32 units, with a minimum of 5 units and a maximum of 186 units. There were 310 patients that received more than 100 units of contrast, while 20 patients received below 50 units.

```{r, echo = FALSE}
par(mfrow = c(1,2))
hist(PE$Contrast.Given, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
#axis(side = 1, at = seq(0,100,10))
boxplot(PE$Contrast.Given, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given")
axis(side = 2, at = seq(0,200,25))
par(mfrow = c(1,1))
summary(PE$Contrast.Given)
```

#### Comparison with Age and BMI

Both the dose of radiation and the amount of contrast given to a patient rely heavily on the patient's age and body composition. As such, we are interested in observing any differences in effective dose and contrast across these two variables. For our purposes, age has been divided into 3 groups: under 18, between 18 and 35, and older than 35. Similarly, BMI had been divided into 3 groups as well: under 25, between 25 and 40, and above 40. Table 1 shows the breakdown of patients that fall into each subgroup.

\begin{table}[H] \centering 
  \caption{No. of Patients by Age and BMI groups} 
\begin{tabular}{p{1cm}p{1cm}|p{3cm}p{3cm}p{3cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
 & & & BMI & \\
 & & $<$25 & 25-40 & $>$40 \\ 
\hline \\[-1.8ex] 
    & $<$18 & 12 (0.4) & 10 (0.3) & 2 (0.1) & 24 (0.9) \\
AGE & 18-35 & 95 (3.6) & 145 (5.4) & 61 (2.3) & 301 (11.3) \\
    & $>$35 & 650 (24.3) & 1336 (50.0) & 363 (13.6) & 2349 (87.8) \\
\hline
    & & 757 (28.3) & 1491 (55.8) & 426 (15.9) & 2674 (100.0) \\
\hline 
\hline \\[-1.8ex]
\multicolumn{6}{l}{\textit{Note:} Percentage of total patients is given in parentheses}
 \end{tabular}
\end{table}

Here we see that over 87% of all patients examined were over 35 years of age. Moreover, approximately half of all patients examined were both over 35 years of age and above a BMI between  25 and 40. This suggests that just looking at effective dose and contrast without accounting for age and BMI will skew our results toward older, overweight but not obese patients. Tables 2 and 3 give the average effective dose and contrast administered, broken down into these same age and BMI categories.

\begin{table}[H] \centering 
  \caption{Average Effective Dose by Age and BMI groups} 
\begin{tabular}{p{1cm}p{1cm}|p{3cm}p{3cm}p{3cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
 & & & BMI & \\
 & & $<$25 & 25-40 & $>$40 \\ 
\hline \\[-1.8ex] 
    & $<$18 & 1.540 & 4.416 & 10.633 & 3.412 \\
AGE & 18-35 & 4.008 & 6.086 & 13.983 & 7.002 \\
    & $>$35 & 4.574 & 7.060 & 13.194 & 7.296 \\
\hline
    &       & 4.455 & 6.948 & 13.295 & 7.227 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

\begin{table}[H] \centering 
  \caption{Average Contrast Given by Age and BMI groups} 
\begin{tabular}{p{1cm}p{1cm}|p{3cm}p{3cm}p{3cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
 & & & BMI & \\
 & & $<$25 & 25-40 & $>$40 \\ 
\hline \\[-1.8ex] 
    & $<$18 & 59.08 & 69.60 & 95.50 & 66.28 \\
AGE & 18-35 & 84.61 & 89.18 & 102.83 & 90.56 \\
    & $>$35 & 87.85 & 90.59 & 103.80 & 91.68 \\
\hline
    &       & 86.98 & 102.94 & 90.31 & 91.32 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

From Table 3, it appears that the effective dose of radiation administered increases with age and BMI, and it seems as if BMI has a greater impact on dose than age (This effect will be concretely determined in the later modeling). In Table 4, we see a drastic increase in the amount of contrast used between the under 18 age group and the 18-35 group. There may be a slight increase between this middle group and patients over 35, but the difference appears to be minimal. There does not seem to be any major difference in contrast used across varying BMI groups.

#### Effect of Scanner Type

The various CT scans were performed using one of four different types of CT scanners, each with different output levels of radiation and varying prerequisites that must be met before using a specific scanner. Table 4 shows the average effective dose and contrast given to patients before a scan on each type of scanner. The "DS_Force" and "DS_Drive" scanners were consolodated into a single variable labelled "DS_Scanner".

\begin{table}[H] \centering 
  \caption{Average Radiation and Contrast by Scanner Type} 
\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
  & 64SSWoIR & 64SSwIR & 128SSwIR & DS Scanner \\ 
\hline \\[-1.8ex] 
 Effective Dose & 10.696 (8.63) & 9.305 (10.87) & 7.113 (6.87) & 6.962 (8.43) \\
 Contrast & 62.41 (43.36) & 67.01 (40.13) & 68.31 (36.48) & 62.72 (33.47)\\
\hline 
\hline \\[-1.8ex]
\multicolumn{5}{l}{\textit{Note:} One standard deviation is given in parentheses}
 \end{tabular}
\end{table}

It appears, from Table 4, that there are not any apparant differences in administered radiation or contrast between the four types of scanners due to the relatively large standard deviations present. However, this is merely the direct effect of scanner type on contrast and radiation dosage. After accounting for other variables during regression, we may see this effect change.

## Positivity Rate

The proportion of tests in which a positive result was achieved in detecting a pulmonary embolism will be referred to as the positivity rate. The overall positivity rate for this data was 10.93%, which corresponds to 296 positive test results. While this is descriptive of the patient base as a whole, the focus of the summary statistics in this section will be on the positivity rate for sub-groups of several key variables.

#### Positivity and Radiation/Contrast

Figure 3a gives the distribution of radiation administered to patients, group by whether or not their CT scan returned positive results. Figure 3b does the same for the distribution of contrast. In either case, a test result of 0 indicates a negative scan. There does not appear to be any substantial difference in the compounds administered to patients when we group by test result. One thing worth noting here is that when looking at the effective dose of radiation administered, we see more outliers in the case where test results ended up being negative. This would suggest that these patients are being unnecessarilly exposed to large amounts of radiation, the top end being over 100 sieverts.

```{r}
PE$PE.Num <- ifelse(PE$Pulmonary.Embolism == "NO", 0, 1)
par(mfrow = c(1,2))
boxplot(PE$Eff.Dose~PE$PE.Num, main = "Figure 3a: Distribution of Effective Dose",
        xlab = "Test Result", ylab = "Effective Dose", cex.main = 0.75)
boxplot(PE$Contrast.Given~PE$PE.Num, main = "Figure 3b: Distribution of Contrast Administered",
        xlab = "Test Result", ylab = "Contrast", cex.main = 0.75)
par(mfrow = c(1,1))
```

#### Positivity and Type of Procedure

The three types of procedures present in the dataset are chest only, double rule out, and chest, abdomen, and pelvis. There was one case of a triple rule out, which has been excluded from analysis. Table 5 provides both the counts and positivity rate for these types of procedures. Although there appears to be a difference in positivity between Chest Only scans (10.9%) and CAP scans (12.3%), this difference was found to be non-significant through a two-proportion Z test (Z = 0.708, p = 0.239). Interestingly, all 43 double rule out procedures produced negative results, which may suggest that double rule out procedures are being over-used.

\begin{table}[H] \centering 
  \caption{Positivity Rate by Type of Procedure} 
\begin{tabular}{p{5cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & n & Positivity \\
\hline \\[-1.8ex] 
Chest Only    & 2369 & 0.109  \\
Chest, Abdomen, and Pelvis    & 301 & 0.123  \\
Double Rule Out    & 43 & 0.000  \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

#### Positivity by Age and BMI

Once again, we are interested in the rate of positive test results after subdividing the patients into the three age and BMI groups used previously. These positivity rates are given in Table 6. From this table it is hard to conclude anything regarding a pattern in positivity rate for Age group or BMI group. There may be a slight decrease in positivity for older age groups, but this may be caused by the relatively small sample size of younger patients.

\begin{table}[H] \centering 
  \caption{Positivity Rate by Age and BMI groups} 
\begin{tabular}{p{1cm}p{1cm}|p{3cm}p{3cm}p{3cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
 & & & BMI & \\
 & & $<$25 & 25-40 & $>$40 \\ 
\hline \\[-1.8ex] 
    & $<$18 & 0.083 & 0.300 & 0.000 & 0.160 \\
AGE & 18-35 & 0.116 & 0.124 & 0.115 & 0.117 \\
    & $>$35 & 0.091 & 0.118 & 0.099 & 0.108 \\
\hline
    &       & 0.094 & 0.119 & 0.101 & 0.109 \\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

#### Positivity and Location of Patient

We are also interested in any difference in positivity dependent on where the patient came from prior to their CT scan. Both the number of patients from each location and the positivity rate are given in Table 7. We can see here that both the ICU and IN locations have a higher rate of positive test results than either the ED or OUT patients. While this appears to be a substantial difference given the sample size for each group, we will determine significance through later regression.

\begin{table}[H] \centering 
  \caption{Positivity Rate by Patient Location} 
\begin{tabular}{p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & n & Positivity \\
\hline \\[-1.8ex] 
ED    & 1537 & 0.096  \\
OUT    & 352 & 0.085  \\
IN    & 613 & 0.124  \\
ICU & 211 & 0.153\\
\hline 
\hline \\[-1.8ex]
 \end{tabular}
\end{table}

## Positive Test Result Statistics

Beyond simply looking at positivity rates, we are also interested in determining patterns within the group of patients that test positive for a pulmonary embolism. More specifically, we will be looking at differences in the type of pulmonary embolism as well as the type of strain detected by the CT scan. Overall, there were 296 patients who tested positive. One patient was recorded as testing positive but did not have information regarding the type of pulmonary embolism or strain, so they were removed from the following summaries, giving us 295 patients for analysis.

#### Type of PE

The number of patients testing positive for each type of pulmonary embolism, divided by location of admittance, is given in Table 8. Here we notice that half of all patients testing positive come from the emergency department category. However, as we saw previously, ED had the lowest positivity rate of all locations, suggesting that the reason for having som many positive cases is that most patients receiving CT scans came from ED. We can also see that a majority of pulmonary embolisms detected are classified as peripheral.

\begin{table}[H] \centering 
  \caption{Type of PE by Location Admitted} 
\begin{tabular}{p{1.5cm}p{3.5cm}|p{2cm}p{2cm}p{2cm}p{2cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
 & & &  \multicolumn{2}{c}{Location} & & \\
 & & ED & OUT & IN & ICU & \\ 
\hline \\[-1.8ex] 
        & Central & 13 (4.4) & 0 (0.0) & 2 (0.7) & 2 (0.7) & 17 (5.8) \\
PE Type & Central and Peripheral & 20 (6.8) & 1 (0.3) & 11 (3.7) & 10 (3.4) & 42 (14.2) \\
        & Peripheral & 115 (39.0) & 17 (5.8) & 62 (21.0) & 42 (14.2) & 236 (80.0) \\
\hline 
    & & 148 (50.2) & 18 (6.1) & 75 (25.8) & 54 (18.3) & 295 (100.0) \\
\hline 
\hline \\[-1.8ex]
\multicolumn{7}{l}{\textit{Note:} Percentage of total patients is given in parentheses}
 \end{tabular}
\end{table}

Table 9 shows the breakdown of type of PE by patient age. We find that a vast majority of positive test results come from patients over 35 years of age. Again however, this is not necessarily due to anything except the fact that most CT scans were performed on patients over 35. This is evidenced by the fact that this age range did not have a larger positivity rate than either of the other two.

\begin{table}[H] \centering 
  \caption{Type of PE by Age Group} 
\begin{tabular}{p{1.5cm}p{3.5cm}|p{2cm}p{2cm}p{2cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
        & & & AGE & & \\
        & & $<$18 & 18-35 & $>$35 &  \\ 
\hline \\[-1.8ex] 
        & Central & 0 (0.0) & 6 (2.0) & 11 (3.7) & 17 (5.8) \\
PE Type & Central and Peripheral & 1 (0.3) & 7 (2.4) & 34 (11.5) & 42 (14.2) \\
        & Peripheral & 3 (1.0) & 23 (7.8) & 210 (71.1) & 236 (80.0) \\
\hline
    & & 4 (1.4) & 36 (12.2) & 256 (86.8) & 295 (100) \\
\hline 
\hline \\[-1.8ex]
\multicolumn{6}{l}{\textit{Note:} Percentage of total sample is given in parentheses}
 \end{tabular}
\end{table}

#### Type of Strain

Similar to the type of pulmonary embolism detected, we are also interested in the distribution of whether a strain was  detected on the CT scan, grouped by location of patient admittance and age group. These data are given in Table 10 and 11, respectively. Here we see that a larger proportion of positive CT scans showed no strain. Again, we have higher incidence of strain in the ED and age over 35 categories due to the larger number of patients overall who fall into these groups.

\begin{table}[H] \centering 
  \caption{Detection of Strain by Location Admitted} 
\begin{tabular}{p{1cm}p{2cm}|p{2cm}p{2cm}p{2cm}p{2cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
        & & & \multicolumn{2}{c}{Location} & & \\
        & & ED & OUT & IN & ICU & \\ 
\hline \\[-1.8ex] 
        & No  & 88 (29.8) & 14 (4.7) & 51 (17.3) & 33 (11.2) & 186 (63.1) \\
Strain  & Yes & 60 (20.3) & 4 (1.4) & 24 (8.1) & 21 (7.1) & 109 (36.9) \\
\hline
    & & 148 (50.2) & 18 (6.1) & 75 (25.4) & 54 (18.3) & 295 (100) \\
\hline 
\hline \\[-1.8ex]
\multicolumn{7}{l}{\textit{Note:} Percentage of total sample is given in parentheses}
 \end{tabular}
\end{table}

\begin{table}[H] \centering 
  \caption{Type of PE by Age Group} 
\begin{tabular}{p{1cm}p{3cm}|p{3cm}p{3cm}p{3cm}|p{2cm}}
\\[-1.8ex] \hline 
\hline
        & & & AGE &  \\
        & & $<$18 & 18-35 & $>$35 &  \\ 
\hline \\[-1.8ex] 
        & No  & 2 (0.7) & 18 (6.1) & 166 (56.3) & 186 (63.1) \\
Strain  & Yes & 2 (0.7) & 18 (6.1) & 89 (30.2) & 109 (36.9) \\
\hline
    &       & 4 (1.4) & 36 (12.2) & 256 (86.8) & 295 (100) \\
\hline 
\hline \\[-1.8ex]
\multicolumn{6}{l}{\textit{Note:} Percentage of total sample is given in parentheses}
 \end{tabular}
\end{table}

## Next Steps

All of these tables demonstrate a summary of how some of our variables of interest interact. It is important to note that no conclusions as to the significance of variable effects or differences can be draw here because no actual statistical tests have been carried out (apart from the two-sample proportions test for positivity in chest only and CAP scans).

The next steps here will be to fit a logistic regression model for positivity using the key variables of interest listed above. Additionally, models for mean radiation and mean constrast will be fit to see which variables have significant effects.

### 2019-03-03 R Files/Updated Model/Update_Analysis.Rmd
Added folders to separate analysis

--- 2019-02-27 R Files/Update_Analysis.Rmd
@@
 ```{r}
+## Creating Age and BMI groups
+age_cuts <- c("<20", "20-30", "30-40", "40-50","50-60", "60-70",
+              "70-80", "80-90", ">90")
+PE$Age10 <- cut(PE$Age, breaks = c(-1,seq(20,90,10), 110), labels = age_cuts)
+
+BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
+              "40-45", "45-50", "50-55", "55-60", ">60")
+PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
+summary(PE$Age10)
+summary(PE$BMI10)
+
 ## Plotting Contrast by Age (10 Year Intervals)
-age_cuts <- c("<10", "10-20", "20-30", "30-40", "40-50","50-60", "60-70",
-              "70-80", "80-90", "90-100", ">100")
-PE$Age10 <- cut(PE$Age, breaks = seq(0,110,10), labels = age_cuts)
-summary(PE$Age10)
-
 contrast_by_age <- tapply(PE$Contrast.Given, PE$Age10, "mean")
@@
 points(contrast_by_age)
-axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)
+axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)
 
@@
 points(rad_by_age)
-axis(1, at = 1:11, labels = age_cuts, cex.axis = .75)
+axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)
 
 ## Plotting Contrast by BMI (5 Unit Intervals)
-BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
-              "40-45", "45-50", "50-55", "55-60", ">60")
-PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
-summary(PE$BMI10)
 
@@
 ```{r}
+## Effective Dosage Model
+
 PE$Age_range <- cut(PE$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
@@
 
+PE$Location2 <- as.factor(ifelse(PE$Location == "ICU", "ICU", "Other"))
+
 effdose1 <- lm(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
+summary(effdose1)
+
+effdose2 <- update(effdose1, .~.-Location)
+anova(effdose1, effdose2)
+
+effdose3 <- lm(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location2, data = PE)
+anova(effdose1, effdose3)
+summary(effdose3)
+
+effdose3.aov <- aov(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location2, data = PE)
+TukeyHSD(effdose3.aov, which = "CT.Type")
 ```
@@
 ```{r}
+## Contrast Model
+
 contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
+summary(contrast1)
+
+contrast2 <- update(contrast1, .~.-Age_range-BMI_range)
+anova(contrast1, contrast2)
+summary(contrast2)
+
+contrast3 <- update(contrast2, .~.-Gender)
+anova(contrast2, contrast3)
+summary(contrast3)
+
+contrast3.aov <- aov(Contrast.Given~CT.Type+Location, data = PE)
+TukeyHSD(contrast3.aov)
 ```

### 2019-03-03 R Files/Updated Model/Updates_Writeup.Rmd
Added folders to separate analysis
Fixed wording

--- 2019-03-01 R Files/Updates_Writeup.Rmd
@@
 Now looking at BMI, there appears to be a linear association between radiation and BMI, with an increase in effective dose as BMI increases.
-Again, there does not appear to be any relationship between contrast and BMI
-
-From these results, it appears that the amount of radiation and contrast administered to a patient are correlated.
-Indeed, we find a Pearson's correlation of 0.47, which is significant (t(2380) = 26.05, p < .001).
+Again, there does not appear to be any relationship between contrast and BMI.
 
@@
 
-As with the effective dose model, post hoc analysis using Tukey's HSD for the type of scanner.
+As with the effective dose model, post hoc analysis was conducted using Tukey's HSD for the type of scanner.
 Here we find that the only significant difference was between the 64SSwIR and 64SSwoIR scanners ($p = .022$).

### 2019-03-03 R Files/Updated Model/Updates_Writeup.tex
Fixed wording

--- 2019-03-01 R Files/Updates_Writeup.tex
@@
 Again, there does not appear to be any relationship between contrast and
-BMI
-
-From these results, it appears that the amount of radiation and contrast
-administered to a patient are correlated. Indeed, we find a Pearson's
-correlation of 0.47, which is significant (t(2380) = 26.05, p
-\textless{} .001).
+BMI.
 
@@
 
-As with the effective dose model, post hoc analysis using Tukey's HSD
-for the type of scanner. Here we find that the only significant
-difference was between the 64SSwIR and 64SSwoIR scanners (\(p = .022\)).
-From our model, this difference equates to the 64SSwIR scanners using
-6.95 units less contrast than the 64SSwoIR scanners.
+As with the effective dose model, post hoc analysis was conducted using
+Tukey's HSD for the type of scanner. Here we find that the only
+significant difference was between the 64SSwIR and 64SSwoIR scanners
+(\(p = .022\)). From our model, this difference equates to the 64SSwIR
+scanners using 6.95 units less contrast than the 64SSwoIR scanners.
 

### 2019-03-04 R Files/Initial Model/PE_analysis .Rmd
Looking at the erroneous contrast values

---
title: "PE_analysis"
author: "Aidan Mullan
date: "2/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Reading in data
PE_oldraw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_old <- na.omit(subset(PE_oldraw[,c(3,4,5,7:13,16,17)], subset = Procedure == "Chest Only"))
```

```{r}
#####-----------------
## Summ Stats for Effective Dose and Contrast
#####-----------------

summary(PE_old$Eff.Dose)
summary(PE_old$Contrast.Given)

## Looking at Contrast alone: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE_old$Contrast.Given, breaks = 20, main = "Figure 2a: Histogram of Contrast Given",
     xlab = "Contrast Given", ylab = "No. Patients", cex.main = 0.75)
boxplot(PE_old$Contrast.Given, main = "Figure 2b: Boxplot of Contrast Given", cex.main = 0.75, 
        ylab = "Contrast Given")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))
summary(PE_old$Contrast.Given)

## Looking at Effective Dose Alone: Histogram and Boxplot
par(mfrow = c(1,2))
hist(PE_old$Eff.Dose, breaks = 20, main = "Figure 1a: Histogram of Effective Dose",
     xlab = "Effective Dose", ylab = "No. Patients", cex.main = 0.75)
axis(side = 1, at = seq(0,100,10))
boxplot(PE_old$Eff.Dose, main = "Figure 1b: Boxplot of Effective Dose", cex.main = 0.75, 
        ylab = "Effective Dose")
axis(side = 2, at = seq(0,100,10))
par(mfrow = c(1,1))

## Comparing Age and BMI groups for Eff Dose and Contrast
table(PE_old$Age_range, PE$BMI_range)

aggregate(Eff.Dose~Age_range+BMI_range, data = PE_old, FUN = "mean")
aggregate(Eff.Dose~Age_range, data = PE_old, FUN = "mean")
aggregate(Eff.Dose~BMI_range, data = PE_old, FUN = "mean")

## Contrast given by age group and BMI group
aggregate(Contrast.Given~Age_range+BMI_range, data = PE_old, FUN = "mean")
aggregate(Contrast.Given~Age_range, data = PE_old, FUN = "mean")
aggregate(Contrast.Given~BMI_range, data = PE_old, FUN = "mean")

## Comparing Scanner Type for Eff Dose and Contrast
aggregate(Eff.Dose~CT.Type, data = PE_old, FUN = "mean")
aggregate(Contrast.Given~CT.Type, data = PE_old, FUN = "mean")
aggregate(Eff.Dose~CT.Type, data = PE_old, FUN = "sd")
aggregate(Contrast.Given~CT.Type, data = PE_old, FUN = "sd")
```

```{r}
#####-------------------
## Summ Stats for Rate of PE
#####-------------------

PE$PE.Num <- ifelse(PE$Pulmonary.Embolism == "NO", 0, 1)
mean(PE$PE.Num)

## Positivity by Procedure Type
aggregate(PE.Num~Procedure, data = PE_old, FUN = "mean")

## Two-proportion Z test
p1 <- 0.1093; p2 <- 0.1229; n1 <- 2369; n2 <- 301
pp <- (p1*n1 + p2*n2) / (n1 + n2)
z <- (p1-p2)/(sqrt(pp*(1-pp)*(1/n1 + 1/n2)))
pnorm(z)

## Positivity by Age/BMI
mean(PE$PE.Num)
aggregate(PE.Num~Age_range, data = PE_old, FUN = "mean")
aggregate(PE.Num~BMI_range, data = PE_old, FUN = "mean")
aggregate(PE.Num~Age_range+BMI_range, data = PE_old, FUN = "mean")

## Positivity by Location
aggregate(PE.Num~Location, data = PE_old, FUN = "mean")
table(PE$Location)
```

```{r}
#####-----------------
## Summ Stats for Patients Testing Positive
####------------------

PE_Pos <- subset(PE_old, subset = PE.Num == 1)

## Type by Location
table(PE_Pos$Type, PE_Pos$Location)
table(PE_Pos$Type)
table(PE_Pos$Location)

## Type by Age
table(PE_Pos$Type, PE_Pos$Age_range)
table(PE_Pos$Age_range)

## Strain by Location
table(PE_Pos$Strain, PE_Pos$Location)
table(PE_Pos$Strain)

## Strain by Age
table(PE_Pos$Strain, PE_Pos$Age_range)
```

```{r}
#####------------------
## Positivity Model
#####------------------

positivity1 <- glm(PE.Num~Gender+Age_range+Eff.Dose+BMI_range+Procedure+CT.Type+Location,
               family = "binomial", data = PE_old)
summary(positivity1)

positivity2 <- update(positivity1, .~.-CT.Type)
summary(positivity2)
anova(positivity1, positivity2, test = "Chisq")

positivity3 <- update(positivity2, .~.-Age_range)
anova(positivity2, positivity3, test = "Chisq")
summary(positivity3)

positivity4 <- update(positivity3, .~.-Procedure)
anova(positivity3, positivity4, test = "Chisq") 
summary(positivity4)

positivity5 <- update(positivity4, .~.-Eff.Dose)
summary(positivity5)
anova(positivity4, positivity5, test = "Chisq")

positivity6 <- update(positivity5, .~.-Gender)
anova(positivity5, positivity6, test = "Chisq")
summary(positivity6)

positivity7 <- update(positivity6, .~.-BMI_range)
anova(positivity7, positivity6, test = "Chisq")
summary(positivity7)
## Final model has only Location
## Decided to re-model due to "double rule out" not having any positive test results
```

```{r}
## Removing double rule out procedure from analysis due to no positive test results
PE_nodub <- subset(PE_old, subset = Procedure != "Double rule out")
pos_nodub1 <- glm(PE.Num~Gender+Age_range+Eff.Dose+BMI_range+Procedure+CT.Type+Location,
               family = "binomial", data = PE_nodub)
summary(pos_nodub1)

pos_nodub2 <- update(pos_nodub1, .~.-Procedure-Age_range-Eff.Dose-CT.Type-Gender)
anova(pos_nodub1, pos_nodub2, test = "Chisq") 
summary(pos_nodub2)
## Final model has BMI and Location

```

```{r}
#####----------------
## Effective Dose Model
#####----------------

effdose1 <- lm(log(Eff.Dose)~CT.Type+BMI_range+Age_range+Gender+Location, data = PE_old)
summary(effdose1)

effdose2 <- update(effdose1, .~.-Location)
anova(effdose1, effdose2)
## Full model is final model
```

```{r}
#####----------------
## Contrast Model
#####----------------

contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE_old)
summary(contrast1)
```

```{r}
#####---------------
## Plotting Contrast by Age and BMI
#####---------------

age_cuts <- c("<20", "20-30", "30-40", "40-50","50-60", "60-70",
              "70-80", "80-90", ">90")
PE_old$Age10 <- cut(PE_old$Age, breaks = c(-1,seq(20,90,10), 110), labels = age_cuts)

BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
              "40-45", "45-50", "50-55", "55-60", ">60")
PE_old$BMI10 <- cut(PE_old$BMI, breaks = c(seq(10,60,5),92), labels = BMI_cuts)
summary(PE_old$Age10)
summary(PE_old$BMI10)

contrast_by_age <- tapply(PE_old$Contrast.Given, PE_old$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Avergae Contrast Administered by Age")
points(contrast_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)

contrast_by_bmi <- tapply(PE_old$Contrast.Given, PE_old$BMI10, "mean")
plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
     main = "Avergae Contrast Administered by BMI")
points(contrast_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)

rad_by_age <- tapply(PE_old$Eff.Dose, PE_old$Age10, "mean")
plot(rad_by_age, type = "l", ylab = "Effective Dose", xlab = "Age Group", xaxt = "n",
     main = "Mean Effective Dose by Age")
points(rad_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .75)

rad_by_bmi <- tapply(PE_old$Eff.Dose, PE_old$BMI10, "mean")
plot(rad_by_bmi, type = "l", ylab = "Effective Dose", xlab = "BMI Group", xaxt = "n",
     main = "Mean Effective Dose by BMI")
points(rad_by_bmi)
axis(1, at = 1:11, labels = BMI_cuts, cex.axis = .75)
```

### 2019-03-08 R Files/Updated Model/Contrast_Model.Rmd
Re-re-fit the contrast model and began writeup

---
title: "Contrast_Model"
author: "Aidan Mullan"
date: "3/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_short <- PE_raw[,c(4,5,8,9,10,11,12)]
PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")

index <- which(PE_chest$Contrast.Given > 150)
PE <- na.omit(PE_chest[-index,])
```

```{r}
PE$Age_range <- cut(PE$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
PE$BMI_range <- cut(PE$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))

contrast1 <- lm(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
summary(contrast1)

contrast2 <- update(contrast1, .~.-Location)
summary(contrast2)
anova(contrast1, contrast2) # p = .026, prefer full model

contrast.aov <- aov(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
TukeyHSD(contrast.aov, which = "CT.Type")
```

### 2019-03-08 R Files/Updated Model/Contrast_Writeup.Rmd
Re-re-fit the contrast model and began writeup

---
title: "Contrast_Writeup"
author: "Aidan Mullan"
date: "3/7/2019"
header-includes:
  - \usepackage{float}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'htb!')
```

## Contrast Distributions

Using the original dataset to start, all patients that were recorded to have received more than 150cc of contrast were removed. This left us with 2365 patients total. Figure 1 shows the distribution of mean contrast administered by age group and BMI group. Here age is grouped by 10-year intervals and BMI is grouped at 5-unit intervals.

```{r, fig.height = 5, fig.width = 10}
PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_short <- PE_raw[,c(4,5,7,8,9,10,11,12)]
PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
index <- which(PE_chest$Contrast.Given > 150)
PE <- na.omit(PE_chest[-index,])

age_cuts <- c("<20", "20-30", "30-40", "40-50","50-60", "60-70",
              "70-80", "80-90", ">90")
PE$Age10 <- cut(PE$Age, breaks = c(-1,seq(20,90,10), 110), labels = age_cuts)
BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
              "40-45", "45-50", "50-55", ">55")
PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,55,5),92), labels = BMI_cuts)

par(mfrow = c(1,2))
contrast_by_age <- tapply(PE$Contrast.Given, PE$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Figure 1b: Mean Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(70,95))
points(contrast_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
axis(2, at = seq(70,95,5))

contrast_by_bmi <- tapply(PE$Contrast.Given, PE$BMI10, "mean")
plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
     main = "Figure 1b: Mean Contrast Administered by BMI", cex.main = 1, yaxt = "n")
points(contrast_by_bmi)
axis(1, at = 1:10, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
axis(2, at = seq(70,120,10))
par(mfrow = c(1,1))
```

Here we see a somewhat quadratic relationship between Age and contrast, which is similar to what we saw for the distribution of radiation administered. We also have a linear association between BMI and contrast, with the amount of contrast being administered to a patient increasing as the patient's BMI increases.

One question that arises is why do we see a quadratic relationship between Age and Contrast? It may be due to some quadratic relationship between age and BMI, where young and old patients tend to weigh less than patients between 30 and 60. Figure 2 overlays the relationship between age and BMI onto the relationship between age and contrast so that we can compare.

In this graph we do see similar relationships with age for BMI and contrast, which seems to support our theory that there is a quadratic relationship between patient age and BMI that can explain the quadratic nature of the relationship between patient age and the amound of contrast administered prior to the CT scan.

```{r, fig.height = 5, fig.width = 8}
par(mar = c(5,5,3,5))
bmi_by_age <- tapply(PE$BMI, PE$Age10, "mean")
plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
     main = "Figure 2: Mean BMI and Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(70,95))
points(contrast_by_age)
axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
axis(2, at = seq(70,95,5))
par(new = T)
plot(bmi_by_age, type = "l", axes = F, lty = 2, xlab = NA, ylab = NA, col = "red")
points(bmi_by_age, col = "red", pch = 16)
axis(side = 4)
mtext(side = 4, line = 3, "BMI")
legend("topright", legend = c("Contrast", "BMI"), lty = c(1,2), col = c("black", "red"))
```

## Contrast Model

 Then, a linear model was fit for contrast considering age (<18, 18-35, >35), BMI (<20, 20-40, >40), Gender, Location of Admission, and Type of CT scanner used. Model coefficients are given in Table 1.

\begin{table}[H] \centering 
  \caption{Positivity Model Coefficients} 
\begin{tabular}{p{3cm}|p{3cm}p{3cm}p{3cm}}
\\[-1.8ex] \hline 
\hline
 & Estimate & Std. Error & p  \\
\hline \\[-1.8ex] 
 Intercept   & 67.332 & 3.490 & <.001 $***$ \\
 BMI 25-40   & 2.847 & 0.742 & <.001 $***$\\
 BMI $>$40   & 16.443 & 1.004 & <.001 $***$\\
 Loc: ICU    & 2.480 & 1.046 & .018 $*$\\
 Loc: IN     & 0.821 & 0.807 & .309 \\
 Loc: OUT    & 2.988 & 1.292 & .021 $.$\\
 CT: 64SSwIR & 4.771 & 1.647 & .004 $**$\\
 CT: 64SSWoIR & 5.752 & 2.143 & .004 $**$ \\
 CT: DS Scanner & $-$9.316 & 0.676 & <.001 $***$\\
 Age: 18-35 & 19.476 & 3.521 & <.001 $***$\\
 Age: $>$35 & 18.577 & 3.411 & <.001 $***$ \\
 Gender: Male & 4.162 & 0.648 & <.001 $***$ \\
\hline 
\hline \\[-1.8ex]
\multicolumn{4}{l}{\textit{Note:} Significance: $.$ < .1, $*$ <.05, $**$ < .01, $***$ < .001}
 \end{tabular}
\end{table}

#### Influence of Scanner Type

Our contrast model found a significant effect for the type of scanner used. Post-hoc comparisons conducted using Tukey's HSD found that the DS scanners were associated with significantly lower levels of contrast than all 3 other scanners ($p<.001$ for all 3). Additionally, the 64SSwIR and 64SSwoIR scanners had significantly higher levels of contrast than the 128SSwIR scanner (64SSwIR: $p = .001$; 64SSwoIR: $p = .016$). There was no significant difference in contrast usage between the 64SSwIR and 64SSwoIR scanners.

From the model, these significant differences can be interpreted as:
\begin{itemize}
\item On average, patients who receive a CT using the DS scanner are administered 9.32cc less contrast than patients on the 128SSwIR scanner, 14.09cc less contrast than patients on the 64SSwIR scanner, and 15.07cc less contrast than patients on the 64SSwoIR scanner
\item On average, patients who receive a CT using the 128SSwIR scanner are administered 4.77cc less contrast than patients on the 64SSwIR scanner and 5.75cc less contrast than patients on the 64SSwoIR scanner.
\end{itemize}

## Radiation and Age

In earlier write-ups, we have seen a quadratic relationship between radiation administered and patient age. There appears to be a peak in effective dosage that occurs somewhere between 30 and 60 years of age. To better pinpoint exactly where this peak occurs, we first subset the data to patients between 25 and 65 years of age. We then create age groups at five-year intervals and retrieve the mean effective dose for each group. These results are given in Figure 3.

```{r}
PE_midage <- subset(PE, subset = Age > 25 & Age < 65)
PE_midage$Age2 <- cut(PE_midage$Age, breaks = seq(25,65,5), labels = seq(30,65,5))
rad_by_age2 <- tapply(PE_midage$Eff.Dose, PE_midage$Age2, "mean")
```

### 2019-03-08 R Files/Updated Model/Contrast_Writeup.tex
Re-re-fit the contrast model and began writeup

--- 2019-03-08 R Files/Updated Model/Contrast_Writeup.Rmd
@@
----
-title: "Contrast_Writeup"
-author: "Aidan Mullan"
-date: "3/7/2019"
-header-includes:
-  - \usepackage{float}
-output: pdf_document
----
+\documentclass[]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Contrast\_Writeup},
+            pdfauthor={Aidan Mullan},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.pos = 'htb!')
-```
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-## Contrast Distributions
+%%% Change title format to be more compact
+\usepackage{titling}
 
-Using the original dataset to start, all patients that were recorded to have received more than 150cc of contrast were removed.
-This left us with 2365 patients total.
-Figure 1 shows the distribution of mean contrast administered by age group and BMI group.
-Here age is grouped by 10-year intervals and BMI is grouped at 5-unit intervals.
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
 
-```{r, fig.height = 5, fig.width = 10}
-PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
-PE_short <- PE_raw[,c(4,5,7,8,9,10,11,12)]
-PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
-index <- which(PE_chest$Contrast.Given > 150)
-PE <- na.omit(PE_chest[-index,])
+\setlength{\droptitle}{-2em}
+  \title{Contrast\_Writeup}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{Aidan Mullan}
+  \preauthor{\centering\large\emph}
+  \postauthor{\par}
+  \predate{\centering\large\emph}
+  \postdate{\par}
+  \date{3/7/2019}
 
-age_cuts <- c("<20", "20-30", "30-40", "40-50","50-60", "60-70",
-              "70-80", "80-90", ">90")
-PE$Age10 <- cut(PE$Age, breaks = c(-1,seq(20,90,10), 110), labels = age_cuts)
-BMI_cuts <- c("10-15", "15-20", "20-25", "25-30","30-35", "35-40",
-              "40-45", "45-50", "50-55", ">55")
-PE$BMI10 <- cut(PE$BMI, breaks = c(seq(10,55,5),92), labels = BMI_cuts)
+\usepackage{float}
 
-par(mfrow = c(1,2))
-contrast_by_age <- tapply(PE$Contrast.Given, PE$Age10, "mean")
-plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
-     main = "Figure 1b: Mean Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(70,95))
-points(contrast_by_age)
-axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
-axis(2, at = seq(70,95,5))
+\begin{document}
+\maketitle
 
-contrast_by_bmi <- tapply(PE$Contrast.Given, PE$BMI10, "mean")
-plot(contrast_by_bmi, type = "l", ylab = "Contrast Given", xlab = "BMI Group", xaxt = "n",
-     main = "Figure 1b: Mean Contrast Administered by BMI", cex.main = 1, yaxt = "n")
-points(contrast_by_bmi)
-axis(1, at = 1:10, labels = BMI_cuts, cex.axis = .7, mgp = c(3,0.5,0))
-axis(2, at = seq(70,120,10))
-par(mfrow = c(1,1))
-```
+\subsection{Contrast Distributions}\label{contrast-distributions}
 
-Here we see a somewhat quadratic relationship between Age and contrast, which is similar to what we saw for the distribution of radiation administered.
-We also have a linear association between BMI and contrast, with the amount of contrast being administered to a patient increasing as the patient's BMI increases.
+Using the original dataset to start, all patients that were recorded to
+have received more than 150cc of contrast were removed. This left us
+with 2365 patients total. Figure 1 shows the distribution of mean
+contrast administered by age group and BMI group. Here age is grouped by
+10-year intervals and BMI is grouped at 5-unit intervals.
 
-One question that arises is why do we see a quadratic relationship between Age and Contrast?
-It may be due to some quadratic relationship between age and BMI, where young and old patients tend to weigh less than patients between 30 and 60.
-Figure 2 overlays the relationship between age and BMI onto the relationship between age and contrast so that we can compare.
+\includegraphics{Contrast_Writeup_files/figure-latex/unnamed-chunk-1-1.pdf}
 
-In this graph we do see similar relationships with age for BMI and contrast, which seems to support our theory that there is a quadratic relationship between patient age and BMI that can explain the quadratic nature of the relationship between patient age and the amound of contrast administered prior to the CT scan.
+Here we see a somewhat quadratic relationship between Age and contrast,
+which is similar to what we saw for the distribution of radiation
+administered. We also have a linear association between BMI and
+contrast, with the amount of contrast being administered to a patient
+increasing as the patient's BMI increases.
 
-```{r, fig.height = 5, fig.width = 8}
-par(mar = c(5,5,3,5))
-bmi_by_age <- tapply(PE$BMI, PE$Age10, "mean")
-plot(contrast_by_age, type = "l", ylab = "Contrast Given", xlab = "Age Group", xaxt = "n",
-     main = "Figure 2: Mean BMI and Contrast Administered by Age", cex.main = 1, yaxt = "n", ylim = c(70,95))
-points(contrast_by_age)
-axis(1, at = 1:9, labels = age_cuts, cex.axis = .7, mgp = c(3,0.5,0))
-axis(2, at = seq(70,95,5))
-par(new = T)
-plot(bmi_by_age, type = "l", axes = F, lty = 2, xlab = NA, ylab = NA, col = "red")
-points(bmi_by_age, col = "red", pch = 16)
-axis(side = 4)
-mtext(side = 4, line = 3, "BMI")
-legend("topright", legend = c("Contrast", "BMI"), lty = c(1,2), col = c("black", "red"))
-```
+One question that arises is why do we see a quadratic relationship
+between Age and Contrast? It may be due to some quadratic relationship
+between age and BMI, where young and old patients tend to weigh less
+than patients between 30 and 60. Figure 2 overlays the relationship
+between age and BMI onto the relationship between age and contrast so
+that we can compare.
 
-## Contrast Model
+In this graph we do see similar relationships with age for BMI and
+contrast, which seems to support our theory that there is a quadratic
+relationship between patient age and BMI that can explain the quadratic
+nature of the relationship between patient age and the amound of
+contrast administered prior to the CT scan.
 
- Then, a linear model was fit for contrast considering age (<18, 18-35, >35), BMI (<20, 20-40, >40), Gender, Location of Admission, and Type of CT scanner used.
-Model coefficients are given in Table 1.
+\includegraphics{Contrast_Writeup_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+\subsection{Contrast Model}\label{contrast-model}
+
+Then, a linear model was fit for contrast considering age
+(\textless{}18, 18-35, \textgreater{}35), BMI (\textless{}20, 20-40,
+\textgreater{}40), Gender, Location of Admission, and Type of CT scanner
+used. Model coefficients are given in Table 1.
 
@@
 
-#### Influence of Scanner Type
+\paragraph{Influence of Scanner Type}\label{influence-of-scanner-type}
 
-Our contrast model found a significant effect for the type of scanner used.
-Post-hoc comparisons conducted using Tukey's HSD found that the DS scanners were associated with significantly lower levels of contrast than all 3 other scanners ($p<.001$ for all 3).
-Additionally, the 64SSwIR and 64SSwoIR scanners had significantly higher levels of contrast than the 128SSwIR scanner (64SSwIR: $p = .001$; 64SSwoIR: $p = .016$).
-There was no significant difference in contrast usage between the 64SSwIR and 64SSwoIR scanners.
+Our contrast model found a significant effect for the type of scanner
+used. Post-hoc comparisons conducted using Tukey's HSD found that the DS
+scanners were associated with significantly lower levels of contrast
+than all 3 other scanners (\(p<.001\) for all 3). Additionally, the
+64SSwIR and 64SSwoIR scanners had significantly higher levels of
+contrast than the 128SSwIR scanner (64SSwIR: \(p = .001\); 64SSwoIR:
+\(p = .016\)). There was no significant difference in contrast usage
+between the 64SSwIR and 64SSwoIR scanners.
 
 From the model, these significant differences can be interpreted as:
+
 \begin{itemize}
@@
 
-## Radiation and Age
-
-In earlier write-ups, we have seen a quadratic relationship between radiation administered and patient age.
-There appears to be a peak in effective dosage that occurs somewhere between 30 and 60 years of age.
-To better pinpoint exactly where this peak occurs, we first subset the data to patients between 25 and 65 years of age.
-We then create age groups at five-year intervals and retrieve the mean effective dose for each group.
-These results are given in Figure 3.
-
-```{r}
-PE_midage <- subset(PE, subset = Age > 25 & Age < 65)
-PE_midage$Age2 <- cut(PE_midage$Age, breaks = seq(25,65,5), labels = seq(30,65,5))
-rad_by_age2 <- tapply(PE_midage$Eff.Dose, PE_midage$Age2, "mean")
-```
+\end{document}

### 2019-03-08 R Files/Updated Model/Contrast_Writeup.Rmd
Finished contrast model and created write-up

@@
 Using the original dataset to start, all patients that were recorded to have received more than 150cc of contrast were removed.
-This left us with 2365 patients total.
+This left us with 2328 patients total.
 Figure 1 shows the distribution of mean contrast administered by age group and BMI group.
@@
 PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
-PE_short <- PE_raw[,c(4,5,7,8,9,10,11,12)]
+PE_short <- PE_raw[,c(4,5,6,7,8,9,10,11,12)]
 PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
@@
 There appears to be a peak in effective dosage that occurs somewhere between 30 and 60 years of age.
-To better pinpoint exactly where this peak occurs, we first subset the data to patients between 25 and 65 years of age.
-We then create age groups at five-year intervals and retrieve the mean effective dose for each group.
-These results are given in Figure 3.
+To better pinpoint exactly where this peak occurs, we can look at the average effective dose administered to patients grouped by age in 1-year intervals (Figure 3a).
+Here we still see the quadratic relationship, but there is a substantial amount of noise in the graph, which makes it rather difficult to identify the exact year at which effective dose reaches a maximum.
 
-```{r}
-PE_midage <- subset(PE, subset = Age > 25 & Age < 65)
-PE_midage$Age2 <- cut(PE_midage$Age, breaks = seq(25,65,5), labels = seq(30,65,5))
-rad_by_age2 <- tapply(PE_midage$Eff.Dose, PE_midage$Age2, "mean")
+To better determine the age at which the maximum radiation is administered, we can smooth this curve to get a better sense of the general pattern which will isolate the maximum effective dose.
+This smoothing process can be seen in Figures 3b.
+From this we get that the maximum average effective dose was administered to patients age 46.
+However, we can see that a more general range for the maximum radiation is for patients between 43 and 49.
+
+```{r, fig.height = 10, fig.width = 10}
+par(mfrow = c(2,2))
+PE$Age_Year <- cut(PE$Age, breaks = c(0,14:93,105), labels = c(1,15:93,105))
+rad_by_age <- tapply(PE$Eff.Dose, PE$Age_Year, "mean")
+plot(rad_by_age, type = "l", main = "Fig 3a: Mean Radiation by Age", xlab = "Age", xaxt = "n",
+     ylab = "Effective Dose", cex.main = 1)
+axis(1, at = c(1,seq(2,length(rad_by_age),5),length(rad_by_age)-1,length(rad_by_age)),
+     labels = c("<15",seq(15,90,5),"93",">93"), cex.axis = 0.8)
+
+rad_smooth <- loess(rad_by_age~c(1:length(rad_by_age)))
+max_rad <- which(predict(rad_smooth) == max(predict(rad_smooth)))
+plot(rad_by_age, type = "l", main = "Fig 3b: Mean Radiation by Age (Smooth)", xlab = "Age", xaxt = "n",
+     ylab = "Effective Dose", cex.main = 1)
+axis(1, at = c(1,seq(2,length(rad_by_age),5),length(rad_by_age)-1,length(rad_by_age)),
+     labels = c("<15",seq(15,90,5),"93",">93"), cex.axis = 0.8)
+lines(predict(rad_smooth), type = "l", lwd = "2")
+abline(v = max_rad, col = "red")
+abline(v = c(max_rad-3, max_rad+3), col = "red", lty = 2)
+legend("topright", legend = c("Age 46", "Age 44-48"), col = "red", lty = c(1,2))
 ```
+
+\newpage
+
+## Patient Population
+
+Several descriptive statistics for the general patient population were requested for the papers that will be written about this data. I have included this information here.
+
+For the total patient population (Chest Only and CAP), there were 2713 patients.
+This breaks down into 1278 male (47.1\%)  and 1435 (52.9\%) female patients.
+For the chest only condition, there were a total of 2328 patients.
+Out of these, 1093 (47.0\%) were male and 1235 (53.\/%) were female.
+
+For the chest only patients, the range of DLP was 7.0 to 3072.0, with a mean of 393.7 and a median of 305.0.
+
+## Manuscript Results Section
+
+This section is a direct re-write of the manuscript that I was sent for the effective dosage and contrast models, with the missing statistics filled in.
+
+#### Effective Dose Model
+
+There were 2369 total patients (mean age 58.1 years; age range 0.2-104.4 years) with 1099 male patients (mean age 59.42; age range 0.2-96.5 years) and 1243 female patients (mean age 56.9; age range 0.6-104.4 years).
+The mean effective radiation dose was  5.512 mSv (median dose 4.27 mSv; dose range 0.1-43.0 mSv).
+A linear regression model was used to determine which factors were significant indicators for the amount of radiation administered to a patient.
+Since the distribution of effective dose was right skewed, the linear model was fit to the log-transform of effective dose.
+
+The model found significant with patient's age group (<18 years, 18-35 years, >35 years), BMI group (BMI under 20, between 20-40, above 40), gender, location of patient admittance, and type of scanner used for CTA.
+All post-hoc pairwise comparisons were then performed using Tukey's HSD with a FDR correction to compensate for the large number of factor levels being compared.
+
+Mean radiation dose for the under 18 age group was 3.18 mSv, which was significantly lower than the average dose for the 18-35 group (mean: 5.24 mSv, p < .001) and over 35 group (mean: 5.57 mSv, p < .001).
+There was no significant difference between the 18-35 and over 35 groups.
+Patient body habitus was a strong predictor for radiation dose with a mean dose of 3.21 mSv for the normal weight group [BMI <25], which was significantly lower than the overweight group [BMI 25-40] (mean: 5.28 mSv, p < .001) and morbidly obese  group [BMI >40] (mean: 10.25 mSv, p < .001).
+The mean radiation dose administered to the morbidly obese group was also significantly higher than the overweight group (p < .001).
+
+For the location at which the patient was admitted, intensive care unit (ICU) patients were associated with the highest radiation dose at an average of 6.16 mSv, compared to an average of 5.43 mSv for all other locations (p < .001).
+For the type of scanner used, the DS scanners had an average effective dose of 4.90 mSv, which was significantly lower than the 64SSwIR scanner (mean: 6.08 mSv, p < .001), 128SSwIR scanner (mean: 5.94 mSv, p < .001), and 64SSwoIR scanner (mean: 9.29 mSv, p < .001).
+Additionally, the 64SSwoIR scanner was associated with a significantly higher amount of radiation than all other scanners (p < .001 for all).
+There was no difference in administered radiation between the 64SSwIR and 128SSwIR scanners.
+
+\newpage
+
+#### Contrast Model
+
+Analysis of the amount of contrast administered to patients used the same patient base as in the effective dose study.
+The mean contrast dose was 88.48 mL (median dose 95.0 mL; dose range 5.0-150.0 mL).
+A linear regression model was used to determine which factors were significant indicators for the amount of radiation administered to a patient.
+
+The model found significant with patient's age group (<18 years, 18-35 years, >35 years), BMI group (BMI under 20, between 20-40, above 40), gender, location of patient admittance, and type of scanner used for CTA.
+All post-hoc pairwise comparisons were then performed using Tukey's HSD with a FDR correction to compensate for the large number of factor levels being compared.
+
+Mean contrast dose for the under 18 age group was 65.67 mL, which was significantly lower than the average dose for the 18-35 group (mean: 88.29 mL, p < .001) and over 35 group (mean: 88.74 mL, p < .001).
+There was no significant difference between the 18-35 and over 35 groups.
+Patient body habitus was a strong predictor for the amount of contrast administered with a mean dose of 84.46 mL for the normal weight group [BMI <25], which was significantly lower than the overweight group [BMI 25-40] (mean: 87.23 mL, p < .001) and morbidly obese  group [BMI >40] (mean: 99.68 mL, p < .001).
+The mean radiation dose administered to the morbidly obese group was also significantly higher than the overweight group (p < .001).
+
+For the location at which the patient was admitted, intensive care unit (ICU) patients were associated with the highest contrast dose at an average of 90.58 mL, compared to an average of 88.21 mL for all other locations.
+Although this effect was found to be significant in the model, the difference was found to be only non-significant during post-hoc analysis with the FDR correction (p = .171).
+For the type of scanner used, the DS scanners had an average contrast dose of 83.45 mL, which was significantly lower than the 128SSwIR scanner (mean: 92.67 mL, p < .001), 64SSwIR scanner (mean: 98.47 mL, p < .001), and 64SSwoIR scanner (mean: 98.84 mL mSv, p < .001).
+Additionally, the 128SSwIR scanner was associated with a significantly lower amount of contrast than the 64SSwIR (p = .006).
+Only marginal significance was found for the difference between the 128SSwIR and 64SSwoIR scanners (p = .064), and no difference was found between the 64SSwIR and 64SSwoIR scanners.
+
+#### Other Details
+
+All analysis was conducted using R statistical software, version 3.5.0.

### 2019-03-08 R Files/Updated Model/Contrast_Writeup.tex
Finished contrast model and created write-up

@@
 have received more than 150cc of contrast were removed. This left us
-with 2365 patients total. Figure 1 shows the distribution of mean
+with 2328 patients total. Figure 1 shows the distribution of mean
 contrast administered by age group and BMI group. Here age is grouped by
@@
 
+\subsection{Radiation and Age}\label{radiation-and-age}
+
+In earlier write-ups, we have seen a quadratic relationship between
+radiation administered and patient age. There appears to be a peak in
+effective dosage that occurs somewhere between 30 and 60 years of age.
+To better pinpoint exactly where this peak occurs, we can look at the
+average effective dose administered to patients grouped by age in 1-year
+intervals (Figure 3a). Here we still see the quadratic relationship, but
+there is a substantial amount of noise in the graph, which makes it
+rather difficult to identify the exact year at which effective dose
+reaches a maximum.
+
+To better determine the age at which the maximum radiation is
+administered, we can smooth this curve to get a better sense of the
+general pattern which will isolate the maximum effective dose. This
+smoothing process can be seen in Figures 3b. From this we get that the
+maximum average effective dose was administered to patients age 46.
+However, we can see that a more general range for the maximum radiation
+is for patients between 43 and 49.
+
+\includegraphics{Contrast_Writeup_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\newpage
+
+\subsection{Patient Population}\label{patient-population}
+
+Several descriptive statistics for the general patient population were
+requested for the papers that will be written about this data. I have
+included this information here.
+
+For the total patient population (Chest Only and CAP), there were 2713
+patients. This breaks down into 1278 male (47.1\%) and 1435 (52.9\%)
+female patients. For the chest only condition, there were a total of
+2328 patients. Out of these, 1093 (47.0\%) were male and 1235 (53./\%)
+were female.
+
+For the chest only patients, the range of DLP was 7.0 to 3072.0, with a
+mean of 393.7 and a median of 305.0.
+
+\subsection{Manuscript Results
+Section}\label{manuscript-results-section}
+
+This section is a direct re-write of the manuscript that I was sent for
+the effective dosage and contrast models, with the missing statistics
+filled in.
+
+\paragraph{Effective Dose Model}\label{effective-dose-model}
+
+There were 2369 total patients (mean age 58.1 years; age range 0.2-104.4
+years) with 1099 male patients (mean age 59.42; age range 0.2-96.5
+years) and 1243 female patients (mean age 56.9; age range 0.6-104.4
+years). The mean effective radiation dose was 5.512 mSv (median dose
+4.27 mSv; dose range 0.1-43.0 mSv). A linear regression model was used
+to determine which factors were significant indicators for the amount of
+radiation administered to a patient. Since the distribution of effective
+dose was right skewed, the linear model was fit to the log-transform of
+effective dose.
+
+The model found significant with patient's age group (\textless{}18
+years, 18-35 years, \textgreater{}35 years), BMI group (BMI under 20,
+between 20-40, above 40), gender, location of patient admittance, and
+type of scanner used for CTA. All post-hoc pairwise comparisons were
+then performed using Tukey's HSD with a FDR correction to compensate for
+the large number of factor levels being compared.
+
+Mean radiation dose for the under 18 age group was 3.18 mSv, which was
+significantly lower than the average dose for the 18-35 group (mean:
+5.24 mSv, p \textless{} .001) and over 35 group (mean: 5.57 mSv, p
+\textless{} .001). There was no significant difference between the 18-35
+and over 35 groups. Patient body habitus was a strong predictor for
+radiation dose with a mean dose of 3.21 mSv for the normal weight group
+{[}BMI \textless{}25{]}, which was significantly lower than the
+overweight group {[}BMI 25-40{]} (mean: 5.28 mSv, p \textless{} .001)
+and morbidly obese group {[}BMI \textgreater{}40{]} (mean: 10.25 mSv, p
+\textless{} .001). The mean radiation dose administered to the morbidly
+obese group was also significantly higher than the overweight group (p
+\textless{} .001).
+
+For the location at which the patient was admitted, intensive care unit
+(ICU) patients were associated with the highest radiation dose at an
+average of 6.16 mSv, compared to an average of 5.43 mSv for all other
+locations (p \textless{} .001). For the type of scanner used, the DS
+scanners had an average effective dose of 4.90 mSv, which was
+significantly lower than the 64SSwIR scanner (mean: 6.08 mSv, p
+\textless{} .001), 128SSwIR scanner (mean: 5.94 mSv, p \textless{}
+.001), and 64SSwoIR scanner (mean: 9.29 mSv, p \textless{} .001).
+Additionally, the 64SSwoIR scanner was associated with a significantly
+higher amount of radiation than all other scanners (p \textless{} .001
+for all). There was no difference in administered radiation between the
+64SSwIR and 128SSwIR scanners.
+
+\newpage
+
+\paragraph{Contrast Model}\label{contrast-model-1}
+
+Analysis of the amount of contrast administered to patients used the
+same patient base as in the effective dose study. The mean contrast dose
+was 88.48 mL (median dose 95.0 mL; dose range 5.0-150.0 mL). A linear
+regression model was used to determine which factors were significant
+indicators for the amount of radiation administered to a patient.
+
+The model found significant with patient's age group (\textless{}18
+years, 18-35 years, \textgreater{}35 years), BMI group (BMI under 20,
+between 20-40, above 40), gender, location of patient admittance, and
+type of scanner used for CTA. All post-hoc pairwise comparisons were
+then performed using Tukey's HSD with a FDR correction to compensate for
+the large number of factor levels being compared.
+
+Mean contrast dose for the under 18 age group was 65.67 mL, which was
+significantly lower than the average dose for the 18-35 group (mean:
+88.29 mL, p \textless{} .001) and over 35 group (mean: 88.74 mL, p
+\textless{} .001). There was no significant difference between the 18-35
+and over 35 groups. Patient body habitus was a strong predictor for the
+amount of contrast administered with a mean dose of 84.46 mL for the
+normal weight group {[}BMI \textless{}25{]}, which was significantly
+lower than the overweight group {[}BMI 25-40{]} (mean: 87.23 mL, p
+\textless{} .001) and morbidly obese group {[}BMI \textgreater{}40{]}
+(mean: 99.68 mL, p \textless{} .001). The mean radiation dose
+administered to the morbidly obese group was also significantly higher
+than the overweight group (p \textless{} .001).
+
+For the location at which the patient was admitted, intensive care unit
+(ICU) patients were associated with the highest contrast dose at an
+average of 90.58 mL, compared to an average of 88.21 mL for all other
+locations. Although this effect was found to be significant in the
+model, the difference was found to be only non-significant during
+post-hoc analysis with the FDR correction (p = .171). For the type of
+scanner used, the DS scanners had an average contrast dose of 83.45 mL,
+which was significantly lower than the 128SSwIR scanner (mean: 92.67 mL,
+p \textless{} .001), 64SSwIR scanner (mean: 98.47 mL, p \textless{}
+.001), and 64SSwoIR scanner (mean: 98.84 mL mSv, p \textless{} .001).
+Additionally, the 128SSwIR scanner was associated with a significantly
+lower amount of contrast than the 64SSwIR (p = .006). Only marginal
+significance was found for the difference between the 128SSwIR and
+64SSwoIR scanners (p = .064), and no difference was found between the
+64SSwIR and 64SSwoIR scanners.
+
+\paragraph{Other Details}\label{other-details}
+
+All analysis was conducted using R statistical software, version 3.5.0.
+
 \end{document}

### 2019-06-22 R Files/Manuscript_Graphics.Rmd
Created images for manuscript

---
title: "Manuscript_Graphics"
author: "Aidan Mullan"
date: "6/22/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
PE_oldraw <- read.csv("Data/1st_Set/PE_clean.csv")
PE_old <- na.omit(subset(PE_oldraw[,c(3,4,5,7:13,16,17)], subset = Procedure == "Chest Only"))
```

```{r}
## Line Plot of Model Predictions
location <- c("ED", "ED", "ED", "ICU", "ICU", "ICU", "IN", "IN", "IN", "OUT", "OUT", "OUT")
color <- c(rep("grey89", 3), rep("grey69", 3), rep("grey49", 3), rep("grey29", 3))
BMI <- rep(c("<25", "25-40", ">40"), 4)
mod_probs <- c(.082, .107, .091, .133, .171, .146, .107, .139, .118, .072, .094, .079)
probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs, color = color)

pos_preds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location)) +
                geom_line(size = .75) + geom_point(size = 1.5) +
                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
                ggtitle("Predicted CT Positivity by BMI and Admittance Location")

ggsave(filename = "Images/PositivityPreds.jpeg", plot = pos_preds,
       width = 7, height = 5, device = "jpeg")

## Patient Age and BMI Demographics
counts <- as.data.frame(table(PE_oldraw$Age_range, PE_oldraw$BMI_range))
names(counts) <- c("Age", "BMI", "n")
counts$Age <- factor(counts$Age, levels = c("<18", "18-35", ">35"))
counts$BMI <- factor(counts$BMI, levels = c("<25", "25-40", ">40"))

age_bmi <- ggplot(counts, aes(fill = BMI, x = Age, y = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and BMI Group")

ggsave(filename = "Images/AgeBMI.jpeg", plot = age_bmi,
       width = 7, height = 5, device = "jpeg")

```

### 2019-07-18 R Files/Manuscript_Graphics.Rmd
Added more manuscript images

@@
 ```{r}
-## Line Plot of Model Predictions
-location <- c("ED", "ED", "ED", "ICU", "ICU", "ICU", "IN", "IN", "IN", "OUT", "OUT", "OUT")
-color <- c(rep("grey89", 3), rep("grey69", 3), rep("grey49", 3), rep("grey29", 3))
-BMI <- rep(c("<25", "25-40", ">40"), 4)
-mod_probs <- c(.082, .107, .091, .133, .171, .146, .107, .139, .118, .072, .094, .079)
-probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs, color = color)
-
-pos_preds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location)) +
-                geom_line(size = .75) + geom_point(size = 1.5) +
-                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
-                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
-                ggtitle("Predicted CT Positivity by BMI and Admittance Location")
-
-ggsave(filename = "Images/PositivityPreds.jpeg", plot = pos_preds,
-       width = 7, height = 5, device = "jpeg")
 
@@
 
-age_bmi <- ggplot(counts, aes(fill = BMI, x = Age, y = n)) + 
+age_bmi <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
               geom_bar(position = "dodge", stat = "identity") +
+              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
+              scale_y_continuous(breaks = seq(0,1500,250)) + 
+              ylab("Number of Patients") + 
+              ggtitle("Patient Population by Age and BMI Group")
+
+bar_AgeBMI <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
+              geom_bar(position = "dodge", stat = "identity") +
+              geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
               scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
@@
        width = 7, height = 5, device = "jpeg")
+ggsave(filename = "Images/Bar_AgeBMI.jpeg", plot = bar_AgeBMI,
+       width = 7, height = 5, device = "jpeg")
+```
 
+```{r}
+#####--------------------
+## Observed Bar Plots
+#####--------------------
+
+PE_old$Positivity <- if_else(PE_old$Pulmonary.Embolism == "YES", 1, 0)
+
+## Positivity X age, bmi
+
+Pos_Age <- PE_old %>% group_by(Age_range) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_Age$Positivity <- round(Pos_Age$Positivity, 3)
+Pos_Age$Age_range <- factor(Pos_Age$Age_range, levels = c("<18", "18-35", ">35"))
+bar_PosAge <- ggplot(Pos_Age, aes(fill = Age_range, x = Age_range, y = Positivity, label = Positivity)) +
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("Age Group") +
+                     labs(fill = "Age Group") +
+                     theme(legend.position = "none") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by Age")
+
+Pos_BMI <- PE_old %>% group_by(BMI_range) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_BMI$Positivity <- round(Pos_BMI$Positivity, 3)
+Pos_BMI$BMI_range <- factor(Pos_BMI$BMI_range, levels = c("<25", "25-40", ">40"))
+bar_PosBMI <- ggplot(Pos_BMI, aes(fill = BMI_range, x = BMI_range, y = Positivity, label = Positivity)) + 
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("BMI Group") +
+                     labs(fill = "BMI Group") +
+                     theme(legend.position = "none") + 
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI")
+
+ggsave(filename = "Images/Bar_PosAge.jpeg", plot = bar_PosAge,
+       width = 7, height = 5, device = "jpeg")
+ggsave(filename = "Images/Bar_PosBMI.jpeg", plot = bar_PosBMI,
+       width = 7, height = 5, device = "jpeg")
+
+## Positivity X age+gender, bmi+gender
+
+Pos_AgeGender <- PE_old %>% group_by(Age_range, Gender) %>% summarise(n = n(),
+                                                                      Positivity = mean(Positivity))
+Pos_AgeGender$Positivity <- round(Pos_AgeGender$Positivity, 3)
+Pos_AgeGender$Age_range <- factor(Pos_AgeGender$Age_range, levels = c("<18", "18-35", ">35"))
+bar_PosAgeGender <- ggplot(Pos_AgeGender, aes(fill = Gender, x = Age_range,
+                                               y = Positivity, label = Positivity)) +
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity") +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("Age Group") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by Age and Gender ")
+
+Pos_BMIGender <- PE_old %>% group_by(BMI_range, Gender) %>% summarise(n = n(),
+                                                                      Positivity = mean(Positivity))
+Pos_BMIGender$Positivity <- round(Pos_BMIGender$Positivity, 3)
+Pos_BMIGender$BMI_range <- factor(Pos_BMIGender$BMI_range, levels = c("<25", "25-40", ">40"))
+bar_PosBMIGender <- ggplot(Pos_BMIGender, aes(fill = Gender, x = BMI_range,
+                                               y = Positivity, label = Positivity)) + 
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity") +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("BMI Group") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")
+
+ggsave(filename = "Images/Bar_PosAgeGender.jpeg", plot = bar_PosAgeGender,
+       width = 7, height = 5, device = "jpeg")
+ggsave(filename = "Images/Bar_PosBMIGender.jpeg", plot = bar_PosBMIGender,
+       width = 7, height = 5, device = "jpeg")
+
+#####-----------------------
+## Positivity X location+bmi
+
+Pos_BMILocation <- PE_old %>% group_by(BMI_range, Location) %>% summarise(n = n(),
+                                                                      Positivity = mean(Positivity))
+Pos_BMILocation$Positivity <- round(Pos_BMILocation$Positivity, 3)
+Pos_BMILocation$BMI_range <- factor(Pos_BMI$BMI_range, levels = c("<25", "25-40", ">40"))
+bar_PosBMILocation <- ggplot(Pos_BMILocation, aes(fill = Location, x = BMI_range,
+                                               y = Positivity, label = Positivity)) + 
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity") +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("BMI Group") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")
+
+ggsave(filename = "Images/Bar_PosBMILocation.jpeg", plot = bar_PosBMILocation,
+       width = 7, height = 5, device = "jpeg")
+
+bar_PosLocationBMI <- ggplot(Pos_BMILocation, aes(fill = BMI_range, x = Location,
+                                               y = Positivity, label = Positivity)) + 
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity") +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("BMI Group") +
+                     labs(fill = "BMI Group") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")
+
+ggsave(filename = "Images/Bar_PosLocationBMI.jpeg", plot = bar_PosLocationBMI,
+       width = 7, height = 5, device = "jpeg")
+
+#####--------------------
+## Observed Scatter Plots
+#####--------------------
+
+## Positivity X raw age
+
+PE_old$Age5 <- cut(PE_old$Age, breaks = seq(0,105, 5), labels = seq(5,105,5), right = FALSE)
+PosAge5 <- PE_old %>% group_by(Age5) %>% summarise(Positivity = mean(Positivity))
+PosAge5$Age5 <- seq(5,100,5)
+
+scat_PosAge <- ggplot(PosAge5, aes(x = Age5, y = Positivity, color = "dodgerblue")) +
+                geom_point(size = 1.5) +
+                geom_smooth(method = "lm", se = FALSE,
+                                size = 0.5, aes(color = "black")) + 
+                geom_label(aes(color = "black"), x = 95, y = 0.08,
+                           label = expression(paste(R^2, " = .010"))) +
+                scale_colour_manual(values = c("black", "dodgerblue")) + 
+                xlab("Age") + ylab("Pulmonary Embolism Positivity Rate") + 
+                theme(legend.position = "none") +
+                ggtitle("Pulmonary Embolism Positivity Rate by Age")
+
+ggsave(filename = "Images/Scat_PosAge.jpeg", plot = scat_PosAge,
+       width = 7, height = 5, device = "jpeg")
+
+## Positivity X raw bmi
+
+PE_old$BMI5 <- cut(PE_old$Age, breaks = seq(10,95, 5), labels = seq(15,95,5), right = FALSE)
+PosBMI5 <- PE_old %>% group_by(BMI5) %>% summarise(Positivity = mean(Positivity))
+PosBMI5$BMI5 <- seq(15,100,5)
+
+scat_PosBMI <- ggplot(PosBMI5, aes(x = BMI5, y = Positivity, color = "mediumblue")) +
+                geom_point(size = 1.5) +
+                geom_smooth(method = "lm", se = FALSE,
+                                size = 0.5, aes(color = "black")) + 
+                geom_label(aes(color = "black"), x = 95, y = 0.08,
+                           label = expression(paste(R^2, " = 0.051"))) +
+                scale_colour_manual(values = c("black", "mediumblue")) + 
+                xlab("BMI") + ylab("Pulmonary Embolism Positivity Rate") + 
+                theme(legend.position = "none") +
+                ggtitle("Pulmonary Embolism Positivity Rate by BMI")
+
+ggsave(filename = "Images/Scat_PosBMI.jpeg", plot = scat_PosBMI,
+       width = 7, height = 5, device = "jpeg")
+
+#####--------------------
+## Observed Bar + Pie
+#####--------------------
+
+## Number patients by location
+
+Pos_Location <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_Location$Positivity <- round(Pos_Location$Positivity, 3)
+
+bar_NumLocation <- ggplot(Pos_Location, aes(fill = Location, x = Location,
+                                       y = n, label = n)) +
+                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
+                     geom_text(size = 3, position = position_dodge(0.5), vjust = -1) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
+                     ylab("Number of Patients") + 
+                     xlab("Patient Location") +
+                     #labs(fill = "Age Group") +
+                     ggtitle("Number of Patients by Admittance Location")
+
+ggsave(filename = "Images/Bar_NumLocation.jpeg", plot = bar_NumLocation,
+       width = 7, height = 5, device = "jpeg")
+
+Pos_Location$pos <- c(189/2, 189+(534/2), 189+534+(268/2), 189+534+268+(1341/2))
+Pos_Location$percent <- paste(round(Pos_Location$n/sum(Pos_Location$n) * 100, digits = 1), "%", sep = "")
+Pos_Location$label <- paste(" ", Pos_Location$Location, " \n ", Pos_Location$percent)
+Pos_Location$label1 <- c(Pos_Location$label[1], "", Pos_Location$label[3:4])
+Pos_Location$label2 <- c("", Pos_Location$label[2], "", "")
+
+pie_NumLocation <- ggplot(Pos_Location, aes(x=factor(1), y=n, fill=Location)) +
+                   geom_bar(stat="identity", color = "black") +
+                   geom_text(aes(x= factor(1), y=rev(pos), label = label1), size=3.5, vjust = -.1) +
+                   geom_text(aes(x= factor(1), y=rev(pos), label = label2), size=3.5) +
+                   scale_fill_manual(values = rev(c("lightskyblue2", "lightblue",
+                                                    "lightskyblue3","lightskyblue4"))) +
+                   theme(axis.ticks = element_blank(), axis.text = element_blank(),
+                         axis.title = element_blank(), panel.background = element_blank(),
+                         legend.position = "none") +
+                   ggtitle("Number of Patients by Admittance Location") +# note y = pos
+                   coord_polar(theta = "y")
+
+ggsave(filename = "Images/pie_NumLocation.jpeg", plot = pie_NumLocation,
+       width = 7, height = 5, device = "jpeg")
 ```
+
+```{r}
+#####----------------------
+## Results Graphics
+#####----------------------
+
+## Line Plot of Model Predictions
+location <- c("ED", "ED", "ED", "ICU", "ICU", "ICU", "IN", "IN", "IN", "OUT", "OUT", "OUT")
+BMI <- rep(c("<25", "25-40", ">40"), 4)
+mod_probs <- c(.082, .107, .091, .133, .171, .146, .107, .139, .118, .072, .094, .079)
+probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs, color = color)
+
+line_PosPreds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location)) +
+                geom_line(size = .75) + geom_point(size = 1.5) +
+                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
+                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
+                ggtitle("Predicted CT Positivity by BMI and Admittance Location")
+
+line_labels_PosPreds <- ggplot(probs, aes(x = bmi, y = probs, group = Location, color = Location,
+                                   label = probs)) +
+                geom_line(size = .75) + geom_point(size = 1.5) +
+                geom_text(size = 3, vjust = -1, color = "black") +
+                scale_colour_manual(values = c("lightblue", "dodgerblue", "mediumblue", "navyblue")) + 
+                xlab("BMI Group") + ylab("Probability of Positive Scan") + 
+                ggtitle("Predicted CT Positivity by BMI and Admittance Location")
+
+ggsave(filename = "Images/Line_PosPreds.jpeg", plot = line_PosPreds,
+       width = 7, height = 5, device = "jpeg")
+ggsave(filename = "Images/Line_Labels_PosPreds.jpeg", plot = line_labels_PosPreds,
+       width = 7, height = 5, device = "jpeg")
+```

### 2019-07-18 R Files/Manuscript_Graphics.Rmd
Image updates

@@
 Pos_BMI$Positivity <- round(Pos_BMI$Positivity, 3)
+Pos_BMI$Signif <- c("*", "", "*")
 Pos_BMI$BMI_range <- factor(Pos_BMI$BMI_range, levels = c("<25", "25-40", ">40"))
@@
                      geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     geom_text(aes(label = Signif), size = 5, position = position_dodge(0.9),
+                               vjust = -1.5) +
                      scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
@@
 Pos_AgeGender$Positivity <- round(Pos_AgeGender$Positivity, 3)
+Pos_AgeGender$Signif <- c("","","","","*","*")
 Pos_AgeGender$Age_range <- factor(Pos_AgeGender$Age_range, levels = c("<18", "18-35", ">35"))
@@
                      geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     geom_text(aes(label = Signif), size = 5, position = position_dodge(0.9),
+                               vjust = -1.5) +
                      scale_fill_manual(values = c("lightblue", "navyblue")) +
@@
                      xlab("BMI Group") +
-                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Location")
 
@@
                      labs(fill = "BMI Group") +
-                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Gender ")
+                     ggtitle("Pulmonary Embolism Positivity Rate by BMI and Location")
 
@@
 mod_probs <- c(.082, .107, .091, .133, .171, .146, .107, .139, .118, .072, .094, .079)
-probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs, color = color)
+probs <- data.frame(Location = location, bmi = BMI, probs = mod_probs)
+probs$bmi <- factor(probs$bmi, levels = c("<25", "25-40", ">40"))
 
@@
        width = 7, height = 5, device = "jpeg")
+
+Pos_Loc <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_Loc$Positivity <- round(Pos_Loc$Positivity, 3)
+Pos_Loc$Signif1 <- c("*", "**", "*", "**")
+Pos_Loc$Signif2 <- c("***", "***", "", "")
+bar_PosLoc <- ggplot(Pos_Loc, aes(fill = Location, x = Location, y = Positivity, label = Positivity)) +
+                     ylim(0,0.25) +
+                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
+                     geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
+                     geom_text(aes(label = Signif1), size = 5, position = position_dodge(0.9), 
+                               vjust = -1.5) +
+                     geom_text(aes(label = Signif2), size = 5, position = position_dodge(0.9),
+                               vjust = -2.5) +
+                     scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue" ,"navyblue")) +
+                     ylab("Pulmonary Embolism Positivity Rate") + 
+                     xlab("Location") +
+                     labs(fill = "Location") +
+                     theme(legend.position = "none") +
+                     ggtitle("Pulmonary Embolism Positivity Rate by Location")
+
+ggsave(filename = "Images/Bar_PosLocation.jpeg", plot = bar_PosLoc,
+       width = 7, height = 5, device = "jpeg")
+
+p1 <- .122; p2 <- .098; n1 <- 1097; n2 <- 1235
+pp <- (p1*n1 + p2*n2) / (n1 + n2)
+z <- (p1-p2)/(sqrt(pp*(1-pp)*(1/n1 + 1/n2)))
+1-pnorm(z)
 ```
+
+## Fix BMI graph
+## Gender positivity (M = .122, F = .098)
+## Gender X Age significance, add asterisks
+--- 18-35 M v. 18-35 F p = .0295, no other significance
+## BMI significance + p-value (chi-square), add asterisks
+--- 25-40 higher than <25, p = .024
+## Positivity by location, add asterisks, provide caption with p-values
+--- ICU > OUT/ED p < .01, IN > ED p = .026
+## Fix title of bmi + location
+## Give caption for BMI scatterplot

### 2019-07-19 R Files/Manuscript_Graphics.Rmd
Updated images

@@
 PE_oldraw <- read.csv("Data/1st_Set/PE_clean.csv")
-PE_old <- na.omit(subset(PE_oldraw[,c(3,4,5,7:13,16,17)], subset = Procedure == "Chest Only"))
+PE_old <- PE_oldraw[,c(3,4,5,7:13,16,17)]
+PE_old$Positivity <- if_else(PE_old$Pulmonary.Embolism == "YES", 1, 0)
+PE_old2 <- subset(PE_old, !is.na(PE_old$BMI))
 ```
@@
 ```{r}
-
 ## Patient Age and BMI Demographics
@@
 
-PE_old$Positivity <- if_else(PE_old$Pulmonary.Embolism == "YES", 1, 0)
-
 ## Positivity X age, bmi
@@
 
-Pos_BMI <- PE_old %>% group_by(BMI_range) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_BMI <- PE_old2 %>% group_by(BMI_range) %>% summarise(n = n(), Positivity = mean(Positivity))
 Pos_BMI$Positivity <- round(Pos_BMI$Positivity, 3)
@@
 
-Pos_BMILocation <- PE_old %>% group_by(BMI_range, Location) %>% summarise(n = n(),
+Pos_BMILocation <- PE_old2 %>% group_by(BMI_range, Location) %>% summarise(n = n(),
                                                                       Positivity = mean(Positivity))
@@
 
-Pos_Loc <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity))
+Pos_Loc <- PE_old %>% group_by(Location) %>% summarise(n = n(), Positivity = mean(Positivity),
+                                                       Cases = sum(Positivity))
 Pos_Loc$Positivity <- round(Pos_Loc$Positivity, 3)
@@
                      geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
-                     geom_text(aes(label = Signif1), size = 5, position = position_dodge(0.9), 
-                               vjust = -1.5) +
-                     geom_text(aes(label = Signif2), size = 5, position = position_dodge(0.9),
-                               vjust = -2.5) +
+                     #geom_text(aes(label = Signif1), size = 5, position = position_dodge(0.9), 
+                               #vjust = -1.5) +
+                     #geom_text(aes(label = Signif2), size = 5, position = position_dodge(0.9),
+                               #vjust = -2.5) +
                      scale_fill_manual(values = c("lightblue", "dodgerblue", "mediumblue" ,"navyblue")) +
@@
 
-p1 <- .122; p2 <- .098; n1 <- 1097; n2 <- 1235
+p1 <- .187; p2 <- .079; n1 <- 107; n2 <- 202
 pp <- (p1*n1 + p2*n2) / (n1 + n2)

### 2019-07-19 R Files/Manuscript_Graphics.Rmd
Cleaning

@@
 ```
-
-## Fix BMI graph
-## Gender positivity (M = .122, F = .098)
-## Gender X Age significance, add asterisks
---- 18-35 M v. 18-35 F p = .0295, no other significance
-## BMI significance + p-value (chi-square), add asterisks
---- 25-40 higher than <25, p = .024
-## Positivity by location, add asterisks, provide caption with p-values
---- ICU > OUT/ED p < .01, IN > ED p = .026
-## Fix title of bmi + location
-## Give caption for BMI scatterplot

### 2019-08-25 R Files/ED_Contrast_Graphics.Rmd
Created images for Eff Dose and Contrast manuscript

---
title: "EffDose_Contrast_Graphics"
author: "Aidan Mullan"
date: "8/25/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(grid)
library(gridExtra)
PE <- read.csv("Data/Updates/PE_update_clean.csv")[,-1]
```

```{r}
## Patient Age and BMI Demographics
PE$Age_range <- cut(PE$Age, breaks = c(0,18,35,110), labels = c("<18", "18-35", ">35"))
PE$BMI_range <- cut(PE$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))
PE$CT.Type <- factor(PE$CT.Type, levels = c("64SSWoIR", "64SSWithIR", "128SSwIR", "DS_Scanner"))
counts <- as.data.frame(table(PE$Age_range, PE$BMI_range))
names(counts) <- c("Age", "BMI", "n")
counts$Age <- factor(counts$Age, levels = c("<18", "18-35", ">35"))
counts$BMI <- factor(counts$BMI, levels = c("<25", "25-40", ">40"))

age_bmi <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and BMI Group")

bar_AgeBMI <- ggplot(counts, aes(fill = BMI, x = Age, y = n, label = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
              scale_fill_manual(values = c("lightblue", "dodgerblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and BMI Group")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Patient_Counts/Bar_AgeBMI_NoLab.jpeg", plot = age_bmi,
       width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Patient_Counts/Bar_AgeBMI.jpeg", plot = bar_AgeBMI,
       width = 7, height = 5, device = "jpeg")

counts2 <- as.data.frame(table(PE$Age_range, PE$Gender))
names(counts2) <- c("Age", "Gender", "n")
counts$Age <- factor(counts$Age, levels = c("<18", "18-35", ">35"))
bar_AgeGender <- ggplot(counts2, aes(fill = Gender, x = Age, y = n, label = n)) + 
              geom_bar(position = "dodge", stat = "identity") +
              geom_text(size = 3, position = position_dodge(0.9), vjust = -1) +
              scale_fill_manual(values = c("lightblue", "navyblue")) +
              scale_y_continuous(breaks = seq(0,1500,250)) + 
              ylab("Number of Patients") + 
              ggtitle("Patient Population by Age and Gender")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Patient_Counts/Bar_AgeGender.jpeg",
       plot = bar_AgeGender, width = 7, height = 5, device = "jpeg")

Location <- PE %>% group_by(Location) %>% summarise(n = n(), Radiation = mean(Contrast.Given),
                                                    Contrast = mean(Contrast.Given))

bar_NumLocation <- ggplot(Location, aes(fill = Location, x = Location,
                                       y = n, label = n)) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.5), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
                     ylab("Number of Patients") + 
                     xlab("Patient Location") +
                     #labs(fill = "Age Group") +
                     ggtitle("Number of Patients by Admittance Location")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Patient_Counts/Bar_Location.jpeg",
       plot = bar_NumLocation, width = 7, height = 5, device = "jpeg")

Scanner <- PE %>% group_by(CT.Type) %>% summarise(n = n(), Radiation = mean(Contrast.Given),
                                                    Contrast = mean(Contrast.Given))

bar_NumScanner <- ggplot(Scanner, aes(fill = CT.Type, x = CT.Type,
                                       y = n, label = n)) +
                     geom_bar(position = "dodge", stat = "identity", width = 0.5) +
                     geom_text(size = 3, position = position_dodge(0.5), vjust = -1) +
                     scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
                     ylab("Number of Patients") + 
                     xlab("Scanner Type") +
                     #labs(fill = "Age Group") +
                     ggtitle("Number of Patients by Type of Scanner Used")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Patient_Counts/Bar_Scanner.jpeg",
       plot = bar_NumScanner, width = 7, height = 5, device = "jpeg")
```

```{r}
#####-------------------
## Effective Dose Figures
#####-------------------
hist_ED <- ggplot(PE, aes(x = Contrast.Given, fill = "dodgerblue", color = "navyblue")) + 
          geom_histogram() +
          scale_fill_manual(values = c("dodgerblue")) +
          scale_color_manual(values = c("navyblue")) +
          xlab("Effective Dose") + ylab("Count") +
          theme(legend.position = "none") +
          ggtitle("Histogram of Effective Dosage")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Hist_ED.jpeg",
       plot = hist_ED, width = 7, height = 4, device = "jpeg")

scat_AgeED <- ggplot(PE, aes(x = Age, y = Contrast.Given, color = "dodgerblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                scale_colour_manual(values = c("black", "dodgerblue")) + 
                xlab("Age") + ylab("Radiation Administered (mSv)") + 
                theme(legend.position = "none") +
                ggtitle("Effective Dosage Administered by Age")

scat_BMIED <- ggplot(PE, aes(x = BMI, y = Contrast.Given, color = "navyblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                scale_colour_manual(values = c("black", "mediumblue")) + 
                xlab("BMI") + ylab("Radiation Administered (mSv)") + 
                theme(legend.position = "none") +
                ggtitle("Effective Dosage Administered by Patient BMI")

combo_scat <- grid.arrange(scat_AgeED, scat_BMIED)

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Scat_AgeED.jpeg",
       plot = scat_AgeED, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Scat_BMIED.jpeg",
       plot = scat_BMIED, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Scat_combo.jpeg",
       plot = grid.arrange(scat_AgeED, scat_BMIED), width = 7, height = 7, device = "jpeg")

box_ScannerED <- ggplot(PE, aes(x = CT.Type, y = Contrast.Given, fill = CT.Type)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
                ylab("Effective Dose (mSv)") + 
                xlab("Patient Location") +
                theme(legend.position = "none") +
                ggtitle("Effective Dose Administered by Location of Admittance")

box_Gender <- ggplot(PE, aes(x = Gender, y = Contrast.Given, fill = Gender)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue","dodgerblue")) +
                ylab("Effective Dose (mSv)") + 
                xlab("Patient Gender") +
                theme(legend.position = "none") +
                ggtitle("Effective Dose Administered by Patient Gender")

box_Age <- ggplot(PE, aes(x = Age_range, y = Contrast.Given, fill = Age_range)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue")) +
                ylab("Effective Dose (mSv)") + 
                xlab("Patient Gender") +
                theme(legend.position = "none") +
                ggtitle("Effective Dose Administered by Patient Gender")

box_BMI <- ggplot(PE, aes(x = BMI_range, y = Contrast.Given, fill = BMI_range)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue")) +
                ylab("Effective Dose (mSv)") + 
                xlab("Patient Gender") +
                theme(legend.position = "none") +
                ggtitle("Effective Dose Administered by Patient Gender")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Box_Scanner.jpeg",
       plot = box_ScannerED, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Box_Gender.jpeg",
       plot = box_Gender, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Box_Age.jpeg",
       plot = box_Age, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Box_BMI.jpeg",
       plot = box_BMI, width = 7, height = 5, device = "jpeg")
```

```{r}
#####-------------------
## Contrast Figures
#####-------------------
hist_Cont <- ggplot(PE, aes(x = Contrast.Given, fill = "dodgerblue", color = "navyblue")) + 
          geom_histogram() +
          scale_fill_manual(values = c("dodgerblue")) +
          scale_color_manual(values = c("navyblue")) +
          xlab("Contrast") + ylab("Count") +
          theme(legend.position = "none") +
          ggtitle("Histogram of Contrast")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Hist_Cont.jpeg",
       plot = hist_Cont, width = 7, height = 4, device = "jpeg")

scat_AgeContrast <- ggplot(PE, aes(x = Age, y = Contrast.Given, color = "dodgerblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                scale_colour_manual(values = c("black", "dodgerblue")) + 
                xlab("Age") + ylab("Contrast Administered (mL)") + 
                theme(legend.position = "none") +
                ggtitle("Contrast Administered by Age")

scat_BMIContrast <- ggplot(PE, aes(x = BMI, y = Contrast.Given, color = "navyblue")) +
                geom_point(size = 1.5) +
                geom_smooth(method = "lm", se = FALSE,
                                size = 0.5, aes(color = "black")) + 
                scale_colour_manual(values = c("black", "mediumblue")) + 
                xlab("BMI") + ylab("Contrast Administered (mSv)") + 
                theme(legend.position = "none") +
                ggtitle("Contrast Administered by Patient BMI")

combo_scat <- grid.arrange(scat_AgeED, scat_BMIED)

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Scat_AgeContrast.jpeg",
       plot = scat_AgeContrast, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Scat_BMIContrast.jpeg",
       plot = scat_BMIContrast, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Scat_combo.jpeg",
       plot = grid.arrange(scat_AgeContrast, scat_BMIContrast), width = 7, height = 7, device = "jpeg")

box_ScannerContrast <- ggplot(PE, aes(x = CT.Type, y = Contrast.Given, fill = CT.Type)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue", "dodgerblue" ,"mediumblue" ,"navyblue")) +
                ylab("Contrast (mL)") + 
                xlab("Scanner Type") +
                theme(legend.position = "none") +
                ggtitle("Contrast Administered by Type of Scanner Used")

box_Location <- ggplot(PE, aes(x = Location, y = Contrast.Given, fill = Location)) +
                geom_boxplot() +
                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue", "navyblue")) +
                ylab("Contrast (mL)") + 
                xlab("Patient Location") +
                theme(legend.position = "none") +
                ggtitle("Contrast Administered by Patient Location of Admittance")

ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Scanner.jpeg",
       plot = box_ScannerContrast, width = 7, height = 5, device = "jpeg")
ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Location.jpeg",
       plot = box_Location, width = 7, height = 5, device = "jpeg")

```

```{r}
p1 <- .187; p2 <- .079; n1 <- 107; n2 <- 202
pp <- (p1*n1 + p2*n2) / (n1 + n2)
z <- (p1-p2)/(sqrt(pp*(1-pp)*(1/n1 + 1/n2)))
1-pnorm(z)
```

### 2019-08-27 R Files/ED_Contrast_Graphics.Rmd
Made images

@@
 PE$BMI_range <- cut(PE$BMI, breaks = c(0,25,40,100), labels = c("<25", "25-40", ">40"))
+
 PE$CT.Type <- factor(PE$CT.Type, levels = c("64SSWoIR", "64SSWithIR", "128SSwIR", "DS_Scanner"))
+PE$CT.Type2 <- ifelse(PE$CT.Type == "64SSWoIR", "64SSwFBP",
+                     ifelse(PE$CT.Type == "64SSWithIR", "64SSwSAFIRE",
+                            ifelse(PE$CT.Type == "128SSwIR", "128SSwSAFIRE", 
+                                   ifelse(PE$CT.Type == "DS_Scanner", "DSwADMIRE", NA))))
+PE$CT.Type2 <- factor(PE$CT.Type2, levels = c("64SSwFBP", "64SSwSAFIRE", "128SSwSAFIRE", "DSwADMIRE"))
+
 counts <- as.data.frame(table(PE$Age_range, PE$BMI_range))
@@
 
-Scanner <- PE %>% group_by(CT.Type) %>% summarise(n = n(), Radiation = mean(Contrast.Given),
+Scanner <- PE %>% group_by(CT.Type2) %>% summarise(n = n(), Radiation = mean(Contrast.Given),
                                                     Contrast = mean(Contrast.Given))
 
-bar_NumScanner <- ggplot(Scanner, aes(fill = CT.Type, x = CT.Type,
+bar_NumScanner <- ggplot(Scanner, aes(fill = CT.Type2, x = CT.Type2,
                                        y = n, label = n)) +
@@
                      #labs(fill = "Age Group") +
+                     theme(legend.position = "none") +
                      ggtitle("Number of Patients by Type of Scanner Used")
@@
 
-box_ScannerED <- ggplot(PE, aes(x = CT.Type, y = Contrast.Given, fill = CT.Type)) +
+box_ScannerED <- ggplot(PE, aes(x = CT.Type2, y = Eff.Dose, fill = CT.Type)) +
                 geom_boxplot() +
@@
 
-box_Gender <- ggplot(PE, aes(x = Gender, y = Contrast.Given, fill = Gender)) +
+box_Gender <- ggplot(PE, aes(x = Gender, y = Eff.Dose, fill = Gender)) +
                 geom_boxplot() +
@@
 
-box_Age <- ggplot(PE, aes(x = Age_range, y = Contrast.Given, fill = Age_range)) +
+box_Age <- ggplot(PE, aes(x = Age_range, y = Eff.Dose, fill = Age_range)) +
                 geom_boxplot() +
@@
                 ylab("Effective Dose (mSv)") + 
-                xlab("Patient Gender") +
+                xlab("Patient Age") +
                 theme(legend.position = "none") +
-                ggtitle("Effective Dose Administered by Patient Gender")
+                ggtitle("Effective Dose Administered by Patient Age Group")
 
-box_BMI <- ggplot(PE, aes(x = BMI_range, y = Contrast.Given, fill = BMI_range)) +
+box_BMI <- ggplot(PE, aes(x = BMI_range, y = Eff.Dose, fill = BMI_range)) +
                 geom_boxplot() +
@@
                 ylab("Effective Dose (mSv)") + 
-                xlab("Patient Gender") +
+                xlab("Patient BMI") +
                 theme(legend.position = "none") +
-                ggtitle("Effective Dose Administered by Patient Gender")
+                ggtitle("Effective Dose Administered by Patient BMI Group")
+
+box_LocationED <- ggplot(PE, aes(x = Location, y = Eff.Dose, fill = Location)) +
+                geom_boxplot() +
+                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue", "navyblue")) +
+                ylab("Effective Dose (mSv)") + 
+                xlab("Patient Location") +
+                theme(legend.position = "none") +
+                ggtitle("Effective Dose Administered by Patient Location")
 
@@
        plot = box_BMI, width = 7, height = 5, device = "jpeg")
+ggsave(filename = "EffDose_Contrast_Manuscript/Images/Eff_Dose/Box_LocationED.jpeg",
+       plot = box_LocationED, width = 7, height = 5, device = "jpeg")
 ```
@@
 
-box_Location <- ggplot(PE, aes(x = Location, y = Contrast.Given, fill = Location)) +
+box_LocationContrast <- ggplot(PE, aes(x = Location, y = Contrast.Given, fill = Location)) +
                 geom_boxplot() +
@@
 
+box_GenderContrast <- ggplot(PE, aes(x = Gender, y = Contrast.Given, fill = Gender)) +
+                geom_boxplot() +
+                scale_fill_manual(values = c("lightblue","dodgerblue")) +
+                ylab("Contrast (mL)") + 
+                xlab("Patient Gender") +
+                theme(legend.position = "none") +
+                ggtitle("Contrast Administered by Patient Gender")
+
+box_AgeContrast <- ggplot(PE, aes(x = Age_range, y = Contrast.Given, fill = Age_range)) +
+                geom_boxplot() +
+                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue")) +
+                ylab("Contrast (mL)") + 
+                xlab("Patient Age") +
+                theme(legend.position = "none") +
+                ggtitle("Contrast Administered by Patient Age Group")
+
+box_BMIContrast <- ggplot(PE, aes(x = BMI_range, y = Contrast.Given, fill = BMI_range)) +
+                geom_boxplot() +
+                scale_fill_manual(values = c("lightblue","dodgerblue", "mediumblue")) +
+                ylab("Contrast (mL)") + 
+                xlab("Patient BMI") +
+                theme(legend.position = "none") +
+                ggtitle("Contrast Administered by Patient BMI Group")
+
 ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Scanner.jpeg",
@@
 ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Location.jpeg",
-       plot = box_Location, width = 7, height = 5, device = "jpeg")
+       plot = box_LocationContrast, width = 7, height = 5, device = "jpeg")
+ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Gender.jpeg",
+       plot = box_GenderContrast, width = 7, height = 5, device = "jpeg")
+ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_Age.jpeg",
+       plot = box_AgeContrast, width = 7, height = 5, device = "jpeg")
+ggsave(filename = "EffDose_Contrast_Manuscript/Images/Contrast/Box_BMI.jpeg",
+       plot = box_BMIContrast, width = 7, height = 5, device = "jpeg")
 

### 2019-08-27 R Files/Updated Model/Contrast_Model.Rmd
Made images

@@
 PE_raw <- read.csv("Data/1st_Set/PE_clean.csv")
-PE_short <- PE_raw[,c(4,5,8,9,10,11,12)]
+PE_short <- PE_raw[,c(4,5,7,8,9,10,11,12)]
 PE_chest <- subset(PE_short, subset = Procedure == "Chest Only")
@@
 contrast.aov <- aov(Contrast.Given~CT.Type+BMI_range+Age_range+Gender+Location, data = PE)
-TukeyHSD(contrast.aov, which = "CT.Type")
+TukeyHSD(contrast.aov)
 ```

### 2019-08-27 R Files/ED_Contrast_Graphics.Rmd
edits

@@
 
-box_ScannerContrast <- ggplot(PE, aes(x = CT.Type, y = Contrast.Given, fill = CT.Type)) +
+box_ScannerContrast <- ggplot(PE, aes(x = CT.Type2, y = Contrast.Given, fill = CT.Type)) +
                 geom_boxplot() +
